---
title: "Lab 5: Factors in Visualizations"
format: 
  html: 
    theme: minty
    fontsize: 1em
    mainfont: sans-serif
    number-sections: true
    number-depth: 2
    code-block-bg: "#76b5c5"
    highlight-style: monochrome
editor: visual
execute: 
  warning: false
  message: false
  eval: false
  echo: false
---

```{r setup}
#| echo: false
#| eval: true

library(tidyverse)
library(forcats)
library(lubridate)

surveys <- read_csv(
  here::here("2_ggplot-import", 
             "Lab", 
             "surveys.csv")
  )
```

The inspiration for this lab comes from the plotting "best practices" that I've learned over the years. My main inspiration is [Will Chase's 2020 RStudio Conference Presentation](https://www.youtube.com/watch?v=h5cTacaWE6I) -- Glamour of Graphics [(slides)](https://www.williamrchase.com/slides/assets/player/KeynoteDHTMLPlayer.html#0).

::: callout-note
Fun fact! The inspiration for the presentation title came from Dr. Kelly Bodwin!
:::

## Revisiting Lab 2

Let's start with the side-by-side boxplots you created in Week 2, to visualize the distribution of weight within each species (not species ID!).

```{r boxplot-jitter-points}
#| eval: true

surveys |> 
  ggplot(mapping = aes(x = weight, y = species)) + 
  geom_jitter(color = "tomato", alpha = 0.05) + 
  geom_boxplot() + 
  labs(x = "Weight (gm)", 
       y = "Species")
```

As you should expect with a character variable, the boxplots go in alphabetical order. This looks rather jumbled, so let's put our factor skills to work!

**1. Reorder the boxplots so the weights go in descending order.**

::: callout-caution
You are required to use functions from `forcats` to complete this task.
:::

**2. Now that you've reordered, look at your y-axis label. That's long! Let's take Will Chase's advice and move the y-axis label to the top of the plot.**

## Time-Series Plot

This week, we focused on learning skills related to working with dates, but we have yet to make a very common type of plot -- the time-series plot. We'll use this plot to motivate a second type of factor reordering!

**1. Create a visualization how weights vary for each genus over the duration of the study.**

Alright, there are a lot of measurements over time! Let's use our dplyr skills to summarize each year and plot the summaries.

**2. Calculate and plot the mean weight for each year (for each genus).**

This should look much less busy! However, you should notice that the legend still goes in alphabetical order.

**3. Reorder the lines so the weights go in descending order.**

::: callout-caution
You are required to use functions from `forcats` to complete this task.
:::

**4. Now that you've reordered the lines, look at your legend's label. That's terrible! Let's give the legend a new title *and* move the information from the y-axis to the plot title.**

## Caputures over the Week

For our final exploration, we will consider the number of rodents captured throughout the week -- transitioning to visualizations of categorical variables.

**1. Create a visualization of the number of rodents captured each day of the week.**

As you might have expected, the ordering of the days of the week is not what we would like.

**2. Change the order of the day of the week to go Monday through Sunday.**

::: callout-note
You can choose to keep the days named as they are (e.g., Mon, Sun), or you can choose to rename the days to their full names (e.g., Monday, Sunday).
:::

::: callout-warning
Your plot **should not** make people tilt their heads to read it! Be mindful in choosing which variable goes on which axis and if / how you use axis labels.
:::

It should be very clear that there are more rodents captured on the weekend than during the week. But, let's explore if this is still the case if we use a "Weekday" / "Weekend" classification system instead.

**3. Collapse Monday through Friday into a `"Weekday"` level, and collapse Saturday and Sunday into a `"Weekend"` level. Plot the number of rodents captured between the two groups.**

::: callout-warning
Your plot **should not** make people tilt their heads to read it! Be mindful in choosing which variable goes on which axis and if / how you use axis labels.
:::
