---
title: ".hand[What is being plotted?]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include = FALSE}
library(tidyverse)
library(gridExtra)
library(flair)

knitr::opts_chunk$set(fig.align = "center", 
                      message = FALSE, 
                      warning = FALSE)
```

.hand-medium[In a scatterplot, the points *are* the observations in the data!]

```{r, echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center'}
ggplot(data = mpg, 
       mapping = aes(x = displ,
                     y = hwy)) + 
  geom_point()
```

---

.hand-medium[But what about a histogram?]

```{r, echo = FALSE, fig.width = 6, fig.height = 6, message=FALSE}
ggplot(data = mpg, 
       mapping = aes(x = hwy)) + 
  geom_histogram()

```

---

.hand-medium[But what about a boxplot?]

```{r, echo = FALSE, fig.width = 6, fig.height = 6, message=FALSE}
ggplot(data = mpg, 
       mapping = aes(x = hwy)) + 
  geom_boxplot()

```

---

.hand-medium[But what about a bar plot?]

```{r, echo = FALSE, fig.width = 6, fig.height = 6, message=FALSE}
ggplot(data = mpg, 
       mapping = aes(x = class)) + 
  geom_bar()

```

---

.larger[Statistical transformations]

.larger[
.pull-left[


- Each of these plots **does not** plot the *raw* data. 

- Instead, they plot some transformation of the data. 

- Each `geom` uses a specific `stat` to calculate the summaries that are plotted.
]
]

.large[
.pull-right[

</br>
</br>

- `geom_histogram()` uses `stat_bin()`

- `geom_boxplot()` uses `stat_boxplot()`

- `geom_bar()` uses `stat_count()`

- `geom_smooth()` uses `stat_smooth()`
]
]
