---
title: "Practice Activity 2: Using Data Visualization to Find the Group"
format: html
editor: visual
---

```{r}
#| include: false
library(tidyverse)
```

Today you will be exploring different types of visualizations to uncover what
months have the longest delays at New York City's airports. 

Some advice:

* Work with those around you

* If you aren't sure how to make a specific plot or how to customize a plot, 
look over the *R for Data Science* textbook, all of the plots in this activity 
are included in the textbook. 

* Google is your friend! If you aren't sure how to accomplish a certain task, 
type what you are trying to accomplish into Google and see what other people are
saying. Generally, adding `ggplot` to the end of your search will help make your
search results more relevant! 

* Finally, don't forget to use your team members and course Discord as
resources. 

# Getting Started

We will be creating visualizations using the **ggplot2** package. Typically, 
you would only need to load in this package at the beginning of your document, 
however the code I've provided you also uses a tool from the **dplyr** package.
So, it would be best for you to load in the **tidyverse** package at the 
beginning of your document. 

For this activity, we will be exploring the `nycflights` data from the
**openintro** package. So, you will need to install the **openintro** package. 

## Creating a Set-up Chunk

1. Insert a code chunk at the beginning of your document (directly under the 
YAML). 
2. Name the code chunk "setup". 
3. Specify that the messages from loading in the packages should not be included 
4. Load in the tidyverse package
5. Load in the openintro package

```{r}

```

# Part One

Create a scatterplot of the relationship between the departure time (`dep_time`)
and the month of the departure (`month`). 

```{r}

```

**Checkpoint:** This plot should be unsatisfying! Think critically about a 
better way to visualize this relationship. 

*Hint:* Think about what type of variable `month` behaves like! 


# Part Two 

In order to create the plot we want, `month` needs to be a character
variable, not a numeric variable. Use the code below to convert `month` into a 
character data type. 

```{r}
#| eval: false
nycflights <- nycflights |> 
  mutate(month = as.character(month))
```


1. Create a visualization of the relationship between the departure delay (`dep_delay`) and the month of the departure (`month`) that allows for you to discern if there are unusual observations with extremely long delays.

```{r}

```

2. Add the origin (airport) of the flight (`origin`) to your visualization. 

```{r}

```


# Part Three

There are two options for how to include `origin` into your visualization. 
However, one method will more easily address the following questions. 

Use your final plot to address the following questions:

1. During what month did the **second** longest delay occur at EWR?  

2. During what month did the **second** longest delay occur at JFK?

3. During what month did the **second** longest delay occur at LGA? 

4. What unit is departure delay measured in?
