---
title: "Practice Activity 2: Data Wrangling with dplyr"
author: "Stat 331"
output: rmdformats::readthedown
---

Today you will be creating and manipulating vectors, lists, and data frames to
uncover a top secret message.

## Dealing with errors

As you work through this PA, you will encounter some code that does not work as
you want it to.  Don't despair!  Errors (when R is unable to run your code) and
bugs (when the code doesn't do what you hoped) are a natural part of coding. 
Even the best of the best deal with these problems regularly - learning to track
down the issue is a skill that you can learn and practice.

### Advice for dealing with errors

**Errors can be sneaky - check results _often_** 

If a chunk of code runs smoothly without giving you any error or warnings this
**does not** necessarily mean it accomplished the desired task.

It is a good habit to check the results of your code **every time** you finish a
task. In general, I would recommend completing the following tasks **every time**
you write a code chunk.

* Include a comment at the beginning of the code chunk that briefly states what
the purpose of the chunk is. Comments in the code chunks come after `#` signs. 
These comments will remind later readers - which might be your future self! - 
what the desired output of the code chunk is.

* If you created a new object, take a look at it! You can inspect the object by
either clicking its name in your *Environment* tab or by typing its name into
the console.  Make sure it looks about how you expect. **Do not** type code to 
inspect the object in your RMarkdown file, as that **is not** code that needs 
to be saved! 

* If you created or updated a data frame, make sure your edits did what you
hoped.  Use the *Environment* or the `head()` and `glimpse()` functions to
investigate your changes.


**Two heads are better than one**

It can be hard to spot bugs in code that you wrote.  

Work with those around you - if something goes wrong, ask a friend or team member
to take a peek at your code and see if any glaring errors (like *syntax error*)
pop out.

**Explain your code out loud**

The best way to troubleshoot a sneaky but is to explain out loud each step of
your code, and what you hoped to accomplish.

Schedule a meeting with your team to have everyone talk through their code. 

If you are alone, try 
[Rubber Duck Debugging](https://en.wikipedia.org/wiki/Rubber_duck_debugging)! 

**Google is your friend**

The whole of the internet is at your disposal!  Use it early, use it often.

Some tricks:

* Copy-paste the exact error message into Google. Chances are, somebody else had a similar problem and got a similar message.

* Include package names in your search terms.  For example, "bar plot in ggplot" is a better search than "bar plot in R".


### Part One: Data import and cleaning

This section will clean today's dataset, so that you can use it more easily in Part Two.


#### Data download & packages

First, we declare our package dependencies and load the data.

(Note that the data loading function `read_csv` will give you an outpouring of helpful information about the dataset.  If you do not see the word "error", there is nothing to be concerned about.)

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)

colleges <- read_csv("https://www.dropbox.com/s/bt5hvctdevhbq6j/colleges.csv?dl=1")
```


#### Data cleaning

Now we will clean the data.  Alas, each of the R chunks in this section will
cause an error and / or do the desired task incorrectly.  (Even the chunks that
run without error **are not** correct!) You will need to find the mistake and
correct it to complete the intended action. 

__Step 1:__ There are too many variables in this dataset.  We don't need all of them.
Narrow your dataset down to only:

- Name of the institution
- City, State, and ZIP code of the institution
- The Admissions Rate
- The average SAT score
- The number of undergraduate students
- The in and out of state tuition
- Whether the school is public or private
- The "REGION" variable.

```{r, eval = FALSE}
colleges_clean <- colleges %>
  select(INSTNM, CITY, STABBR, ZIP,
         ADM_RATE, SAT_AVG, UGDS,
         TUITIONFEE_IN, TUITIONFEE_OUT,
         CONTROL, REGION) 
```

__Step 2:__ Remove the schools that are private and for-profit (category 3).

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>%
  filter(CONTROL == 1, CONTROL == 2) 
```


__Step 4:__ Adjust the appropriate variables to be numeric.


```{r, eval = FALSE}
colleges_clean <- colleges_clean %>%
  mutate(
    TUITIONFEE_IN = numeric(TUITIONFEE_IN),
    TUITIONFEE_OUT = numeric(TUITIONFEE_OUT),
    SAT_AVG = numeric(SAT_AVG),
    ADM_RATE = numeric(ADM_RATE)
    ) 
```

__Step 5:__ Adjust the appropriate variables to be factors.


```{r, eval = FALSE}
colleges_clean %>% mutate(
    CONTROL = as.factor(CONTROL),
    REGION = as.factor(REGION)
)
```

__Step 6:__ Create a new variable called TUITION_DIFF which contains the 
difference between in and out of state costs.

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% 
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
```


__Step 7:__ Remove every row with missing data.  

**_Note_:** This is not always a great idea!  Usually, even if *some* of the
information is missing, we don't want to throw out the entire row.  This time,
however, we'll be lazy.

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% 
  drop.na()
```


Lastly, notice that each of these steps started with 

```{r, eval = FALSE}
colleges_clean <- colleges_clean %>% ...
```

That is pretty redundant!  Instead, we could perform all these tasks as one long
"pipeline."  

__Step 8:__ Combine your (fixed) code chunks into a **single** code chunk that
carries out all of the steps necessary to clean the data.

**Note:** You **should not** have multiple calls to the `mutate()` function! 


## Part Two:  Identify the mystery college

Wow!  Your best friend Ephelia has been accepted to her top choice college!
Unfortunately, Ephelia is a very mysterious person, and she won't tell you
directly which college this is.  You'll have to use her clues to figure out
which school is her dream school.


Clues:

1. This college is located in Region 7.

2. This college's admission rate is above the median rate for the region.

3. This college **does not** charge the same for in- and out-of-state tuition.

4. The average SAT score of this college is an odd number.

5. This college **is not** in Idaho. 

6. Less than 1,000 people apply to this college every year.  (Assume the size of
the freshman class is 1/4 of the undergraduate population.)

7. The college **is not** in the state where Dr. Theobold received a Ph.D. 

8. Of the three options remaining at this step, Ephelia will attend the cheapest
one.

