---
title: "Linear Regression"
format: 
  html: 
    self-contained: true
    theme: minty
    fontsize: 1em
    mainfont: sans-serif
    number-sections: true
    number-depth: 2
    code-block-bg: "#76b5c5"
    highlight-style: monochrome
editor: visual
execute: 
  echo: true
---

```{r setup}
#| include: false

library(tidyverse)
library(palmerpenguins)
```

In this module, you'll learn to compute linear regression models in R.

`r emo::ji("play button")` Watch Videos: 30 minutes

-   `r emo::ji("link")` Link to **all** Slides from Videos: [here](https://app.box.com/s/zcnbbt2xepty8ztg56ikdgdxz8oiyvws)

`r emo::ji("book")` Readings: 45-60 minutes

`r emo::ji("computer")` Activities: 30 minutes

`r emo::ji("check")` Preview Activity: 1

## Review of Simple (Basic) Linear Regression

`r emo::ji("book")` [Recommended Reading -- *Modern Dive* : Basic Regression](https://moderndive.com/5-regression.html)

`r emo::ji("play")` [Recommended Video -- Simple Linear Regression (17 minutes)](https://www.youtube.com/watch?v=1IMpg1ukMpY)

## Conditions of Linear Regression

`r emo::ji("book")` [Recommended Readign -- *Modern Dive*: Conditions for inference for regression](https://moderndive.com/10-inference-for-regression.html#regression-conditions)

`r emo::ji("play")` [Recommended Video -- Regression Assumptions (14 minutes)](https://www.youtube.com/watch?v=LD3OSSVxKMU)

## Visualizing Linear Regression

```{r}
#| message: false
#| warning: false
library(palmerpenguins)

penguins %>%
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  stat_smooth(method = "lm") + 
  labs(x = "Bill Depth (mm)", 
       y = "Bill Length (mm)")
```

## Fitting Linear Regression Models in R -- Supervised Learning

Fitting regression models in R is very simple, you can use the `lm()` function, with the formula: `reponse ~ explanatory`. If you look at the help file for `lm()`, you'll notice that the `data`set is specified in the **second** argument, which makes it so that we need to specify *where* the piped data is being input into the function. 

We can do this "specifying" using the `_` symbol, as seen in the code below:

```{r}
my_model <- penguins |>
  lm(bill_length_mm ~ bill_depth_mm, data = _)
```

## Accessing Model Estimates

### Raw Coefficients

This is what is output when you just call the name of the linear model object you created (`my_model`). Notice, the output doesn't give you much information and it looks kind of bad (not something you'd want to put in a report).

```{r}
my_model
```

### Model Summary

This is what is output when you use the `summary()` function on a linear model object. Notice, the output gives you a lot of information, some of which is really not that useful. And, the output is quite messy!

```{r}
summary(my_model)
```

### Tidy Model Summary

The `tidy()` function from the broom package takes a linear model object and puts the "important" information into a tidy tibble output.

```{r}
broom::tidy(my_model) 
```

Ah! Just right!

If you are sad that you no longer have the statistics about the model fit (e.g., R-squared, adjusted R-squared, $\sigma$), you can use the `glance()` function from the broom package to grab those!

```{r}
broom::glance(my_model)
```

## Preview Activity: Linear Regression

**Question 1:** If you switch the order of `bill_length_mm` and `bill_depth_mm` in the `lm()` formula nothing happens.

- True
- False

**Question 2:** What *object type* was returned by `summary()`?

- list
- vector
- dataframe

**Question 3:** What *object type* was returned by `tidy()`?

- list
- vector
- dataframe


**Question 4:** Penguin Mopsy has a bill that is 5mm deeper than Penguin Gidget. What will be the difference between their bill lengths?

- Mopsy's length is 3.25mm shorter than Gidget
- Mopsy's length is 3.25mm longer than Gidget

**Question 5:**

Run the following code, and explore the results:

```{r}
my_model_2 <- penguins |>
  lm(bill_length_mm ~ bill_depth_mm*species, data = _)
```

Make a plot illustrating `my_model_2`. Upload a picture of your plot! 

::: {.callout-tip}
Hint: Think about what needs to change in the aesthetic of the plot featured previously!
:::

**Question 6:** Which model of the two models explains the most variance in `bill_length_mm`?

- `my_model` 
- `my_model_2`

**Question 7:** Do the three species of penguin have the same bill shape (i.e., the relationship between length and depth)?
