---
title: "Git, GitHub, and RStudio"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
```


**GitHub** is a site for sharing and collaborating on code, and for careful
tracking of changes as you progress.  As of right now, it is the best way to
organize your work in **R**, to showcase it to  employers, or to share it with
collaborators.

I would argue that being able to use a version control system, such as Git, is 
a critical skill for **every** scientist. Using **Git** and **GitHub** takes
some getting used to, and can be frustrating at times even to experts.  

We will start using Git in its most basic form in this course, and leave further 
exploration up to you. 

---

`r emo::ji("play button")` Watch Videos: 45 minutes
  
  * `r emo::ji("link")` Link to **all** Slides from Videos: [here]()

`r emo::ji("book")` Readings: 10-15 minutes   

`r emo::ji("computer")` Activities: 30 minutes  
  
`r emo::ji("check")` Check-ins: 4 (included in 1 assignment on Canvas)

---

## Before We Get Started

[Meet the toolkit: Version control and collaboration](https://www.youtube.com/watch?v=124DQasLyNQ) 

---

## Introduction to Git and GitHub

`r emo::ji("play button")` [Required Video-- Intro to Git and Github](https://www.youtube.com/embed/BCQHnlnPusY)

  * I hope you love this person as much as I do! 

`r emo::ji("question")` [Additional Materials on Git and GitHub](https://swcarpentry.github.io/git-novice/)

---

`r emo::ji("check")` **Check-in: Make a GitHub Account**

Create an account on [GitHub.com](GitHub.com), if you have not already.

Ideally, your username will be related to your real name, e.g. "atheobold"
rather than "SuperKrazy2000".  However, this is not required; choose whatever
username you prefer.  

`r emo::ji("check")` **Check-in: Make a repository** 

Create a **public** repository called "Test Repo" and initialize the repo with 
a README.md file.   

Create a new text file called "My favorite".  Copy-paste your favorite quote or
poem into the file.

Take a screenshot of your repository showing the README.md file and the text 
file you created.  It should look similar to 9:36 of the video.
 
---

## Git and RStudio

`r emo::ji("play button")` [Required Video -- GitHub and RStudio](https://www.youtube.com/embed/qcjpHFwCugE)

`r emo::ji("question")` [Additional Materials on GitHub and RStudio](https://swcarpentry.github.io/git-novice/14-supplemental-rstudio/index.html)

`r emo::ji("play button")` **Check-in: Access GitHub from RStudio** 

Follow the instructions in the video to make a new R Project that is also a 
GitHub Repository.

Take a screenshot showing yourself committing to a new branch. It should look
like XXX of the video.

**Note:** There are a lot of fiddly computer details that might prevent this
from going as smoothly as in the video.  This doesn't necessarily mean you are
doing something wrong!  If you hit a snag, please contact Dr. Theobold to get
help!

---

## Branches and Pull Requests

`r emo::ji("play button")` [Required Video -- Branching](https://www.youtube.com/embed/oPpnCh7InLY)

`r emo::ji("check")` **Check-in: Making a branch and pull request**

Use the `pr_init()` function from the **usethis** package to create a new branch
of your "Test Repo" repository. Name this branch after your favorite animal.

In this branch, replace a word in your text file with the animal's name.

Commit this change. 

Use the `pr_push()` function to create a **pull request** to your `master`
branch from your new branch.

Take a screen shot of your completed pull request.  It should look like XXX of
the video.

---

## Words Matter -- Renaming the `master` branch

`r emo::ji("book")` [Required Reading -- Renaming the master branch](http://www.kapwing.com/blog/how-to-rename-your-master-branch-to-main-in-git/)

There is no Check-In for this reading; I ask only that you read the blog post
and think carefully about the issues.  Please do not skip this section.

If you do feel that you wish to rename your master branch, I am happy to help
you through the process!
