---
title: "Iteration"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(palmerpenguins)
library(lubridate)
```


```{css, echo = FALSE}
img {
  vertical-align:middle;
}

body {
  text-shadow:none;
}

```


```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "30 min", 
                   reading_time = "0 min", 
                   work_time = "60 min", 
                   num_activities = 2)
```

---

```{r, results = "asis", echo = FALSE}
req_vid("Functional Programming", type = "Required", url = "https://www.youtube.com/embed/lq2a4m8TMYo")
```

```{r, results = "asis", echo = FALSE}
req_tutorial("Iteration", type = "Required", url = "https://rstudio.cloud/learn/primers/5")
```

---

```{r, results = "asis", echo = FALSE}
checkin("Functional Programming")
```

Recall that in the last unit, we discussed the challenge of standardizing many
columns in a data frame.  For example, we wanted to standardize all the numeric 
variables in the `penguins` dataset (from the **palmerpenguins** package).

Here is the function we wrote for that:

```{r}
standardize <- function(data) {
  deviations <- data - mean(data, na.rm = TRUE)
  newdata <- deviations / sd(data, na.rm = TRUE)
  return(newdata)
}
```


**Question 1:** Why does the following not work?

```{r}
penguins %>%
  mutate(
    body_mass_g = map_dbl(body_mass_g, standardize)
  )
```

<!-- The `map_dbl()` function expects a single output from a list or atomic vector. -->
<!-- If you use `safely(standardize)` you should get the following error:  -->

<!-- ``` -->
<!-- Error: Result 1 must be a single double, not a list of length 2 -->
<!-- ``` -->

<!-- This means that `map_dbl()` expects that the result should have length 1!  -->


**Question 2:** 

Use appropriate `map` functions, along with `scale()`, to create
a standardized version of the measurement variables from the `penguins` dataset
(`bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`).
Do this **without** copy-pasting lines and changing `bill_length_mm` to
`bill_depth_mm`, etc.

I recommend looking into `map_df()` or `map_at()`! 

```{r, include = FALSE}

penguins %>%
  map_at(vars(bill_length_mm:body_mass_g), 
         standardize)

penguins %>% 
  select(bill_length_mm:body_mass_g) %>% 
  map_df(., standardize)

```

