---
title: "Linear Regression & Simulation"
date: "May 25, 2021"
author: "Stat 331"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---


```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
knitr::opts_chunk$set(fig.align = "center")

options(htmltools.dir.version = FALSE)

library(openintro)
library(broom)
library(gridExtra)
library(kableExtra)
library(png)
library(xaringanthemer)
library(xaringan)
library(tidyverse)
library(flair)
library(moderndive)


style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

class: inverse

.larger[The next two weeks...]

--

.pull-left[
.bitlarger[**Regression & Simulation**]
]

--

.pull.right[

</br>
.bitlarger[**Predictive Modeling**]
]

---

.larger[Linear Regression]

$$
  response = intercept + slope \cdot explanatory + noise
$$
--

.pull-left[

</br>
<center>
$\hat{y} = \hat{\beta}_0 + \hat{\beta}_1 \cdot x$

]

.pull-right[
```{r 2, echo = FALSE, fig.width = 5, fig.height = 4, fig.align = 'center'}
ggplot(data = bdims, aes(y = wgt, x = hgt)) + 
  geom_point() +
  scale_x_continuous("Explanatory Variable", labels = NULL) + 
  scale_y_continuous("Response Variable", labels = NULL)
```
]

---

.larger[Supervised Learning]

```{r}
evals %>% 
  lm(score ~ age, data = .)
```

---

.larger[Accessing Residuals & Fitted Values]

- `resid()` 
- `fitted()`

---

.larger[Plotting Residuals]

```{r, echo = FALSE}
evals_lm <- evals %>% 
  lm(score ~ age, data = .)
```

```{r, out.width = "30%", out.height = "20%"}
ggplot(data = evals_lm, aes(y = .resid, x = .fitted )) +
  geom_point()
```


---

class: inverse

.huge-text[Simulating Data]

---

.larger[Functions in R]

- `r` stands for for **random sampling**: generate a random value from a 
specific distribution

--

- `p` for **probability**: i.e. compute the chance of observing a value **less
than** `x`

--

- `q` stands for **quantile**: given a probability `p`, compute the x-value
such that $P(X < x) = p$

--

- `d` stands for **density**: compute the height of distribution curve

---

.larger[Distributions in R]

- Uniform -- `unif`
- Normal -- `norm`
- t -- `t`
- Chi-Squared -- `chisq`
- Binomial -- `binom`
