---
title: "Linear Regression & Simulation"
format: 
  revealjs:
    self-contained: true
    theme: dark 
editor: visual
execute: 
  echo: false
  message: false
  warning: false
---

```{r set-up}
#| include: false
library(openintro)
library(broom)
library(gridExtra)
library(kableExtra)
library(png)
library(tidyverse)
library(moderndive)
```

##  {background-color="#B6CADA"}

::: {style="font-size: 4em; color: #000000;"}
The next two weeks...
:::

::: columns
::: {.column width="40%"}
::: {style="font-size: 2em; color: #0F4C81;"}
**Regression & Simulation**
:::
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
::: {style="font-size: 2em; color: #0F4C81;"}
**Predictive Modeling**
:::
:::
:::

##

::: {style="font-size: 3em; color: #FFFFFF;"}
Linear Regression
:::

$$
  response = intercept + slope \cdot explanatory + noise
$$

. . .

::: columns
::: {.column width="40%"}
```{r general-scatterplot}
#| fig-width: 5
#| fig-height: 4
#| fig-align: center

ggplot(data = bdims, aes(y = wgt, x = hgt)) + 
  geom_point() +
  scale_x_continuous("Explanatory Variable", labels = NULL) + 
  scale_y_continuous("Response Variable", labels = NULL) + 
  theme(axis.title.x = element_text(size = 18), 
        axis.title.y = element_text(size = 18)
        )
```

:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
</br>
</br>

$\hat{y} = \hat{\beta}_0 + \hat{\beta}_1 \cdot x$

$\epsilon_i \sim N(0, \sigma^2)$
:::
:::

##

::: {style="font-size: 3em; color: #FFFFFF;"}
Fitting a Linear Regression
:::

```{r slr-example}
#| echo: true
#| code-line-numbers: false

lm(score ~ age, data = evals)
```

##

::: {style="font-size: 3em; color: #FFFFFF;"}
Using the `|>`
:::


```{r wrong-dot}
#| eval: false
#| echo: true
#| code-line-numbers: false
evals |>
  lm(score ~ age, data = .)
```

. . .

</br> 

```
Error in is.data.frame(data): object '.' not found
```
. . .

</br>
</br>

<center>
::: {style="font-size: 2em; color: #B6CADA;"}
**What happened???**
:::


##  {background-color="#B6CADA"}

::: {style="font-size: 3em; color: #000000;"}
Placeholder
:::

Piping is an infix operator (e.g., `2 + 2`) that pipes the LHS of the pipe into
the **first** argument of the function on RHS of the pipe

::: columns
::: {.column width="48%"}
```{r dot-notation}
#| echo: true
#| eval: false
#| code-line-numbers: false

evals %>%
  lm(score ~ age,
     data = .)
```

::: {style="font-size: 0.75em; color: #000000;"}
With `%>%`, if the LHS **should not** be used in the **first** argument, a `.`
can be used to reference where the LHS should go.
:::
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
```{r dash-notation}
#| echo: true
#| eval: false
#| code-line-numbers: false

evals |>
  lm(score ~ age, 
     data = _)
```

::: {style="font-size: 0.75em; color: #000000;"}
`|>` **does not** have a placeholder. However, as of R 4.2, `|>` can use `_` as
a placeholder but only with **named** arguments.
:::
:::
:::

[Interesting blog about differences by Isabella Velásquez](https://towardsdatascience.com/understanding-the-native-r-pipe-98dea6d8b61b)

## 

::: {style="font-size: 3em; color: #FFFFFF;"}
Accessing Residuals & Fitted Values
:::

```{r}
evals_lm <- evals |> 
  lm(score ~ age, data = _)
```

::: columns
::: {.column width="35%"}
::: {style="font-size: 0.85em; color: #FFFFFF;"}
***Objects in `evals_lm`***

`$fitted.values`
`$residuals`
:::
:::

::: {.column width="35%"}
::: {style="font-size: 0.85em; color: #FFFFFF;"}
***Functions to extract elements from `evals_lm`***

- `resid()` 
- `fitted()`
:::
:::

::: {.column width="25%"}
::: {style="font-size: 0.85em; color: #FFFFFF;"}
***Function to make a tibble of elements from `evals_lm`***

`broom::augment()`
:::
:::
:::

## 

::: {style="font-size: 3em; color: #FFFFFF;"}
An Example
:::

::: columns
::: {.column width="45%"}
```{r residuals}
#| echo: true
#| code-line-numbers: false
evals_lm$residuals |> 
  head(4)
```

</br>

```{r resid}
#| echo: true
#| code-line-numbers: false
resid(evals_lm) |> 
  head(4)
```
:::

::: {.column width="3%"}
:::

::: {.column width="50%"}
```{r broom}
#| echo: true
#| code-line-numbers: false
broom::augment(evals_lm)
```
:::
:::

## 

::: {style="font-size: 3em; color: #FFFFFF;"}
Plotting Residuals
:::

```{r residuals-plot}
#| echo: true
#| code-line-numbers: false
#| fig-align: center

evals_lm |> 
  broom::augment() |> 
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```


##  {background-color="#B6CADA"}
::: {style="font-size: 5em; #0F4C81;"}
Simulating Data
:::

##

::: {style="font-size: 3em; color: #FFFFFF;"}
Functions in R
:::


- `r` stands for for **random sampling**: generate a random value from a 
specific distribution

. . .

- `p` for **probability** (CDF): i.e. compute the probability of observing a
value **less than** `x` (cumulative)

. . .

- `q` stands for **quantile**: given a probability `p`, compute the x-value
such that $P(X < x) = p$

. . .

- `d` stands for **density** (PDF): compute the height of distribution curve
(absolute)

##

::: {style="font-size: 3em; color: #FFFFFF;"}
Distributions in R
:::

- Uniform -- `unif`

- Normal -- `norm`

- t -- `t`

- Chi-Squared -- `chisq`

- Binomial -- `binom`

