---
title: "Troubleshooting Errors!"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(flair)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

class: center, middle, inverse

.huge[Syntax Errors]

---

class: center, middle

.larger[Did you leave off a parenthesis?]

```{r, echo = FALSE}
decorate('seq(from = 1, to = 10, by = 1', eval = FALSE) %>%
  flair("(")
```

```{r, echo = FALSE, error = TRUE}

seq(from = 1, to = 10, by = 1

```

---

class: center, middle

.larger[Did you leave off a comma?]

```{r, echo = FALSE, error = TRUE}
decorate('seq(from = 1, to = 10 by = 1)', eval = FALSE) %>%
  flair("by")
```

```{r, echo = FALSE, error = TRUE}

seq(from = 1, to = 10 by = 1)

```

---

class: center, middle

.larger[Did you make a typo? Are you using the right names?]


```{r, echo = FALSE}
decorate('sequence(from = 1, to = 10, by = 1)', eval = FALSE) %>%
  flair("sequence")
```

```{r, echo = FALSE, error = TRUE}

sequence(from = 1, to = 10, by = 1)

```

---

class: center, middle

.larger[Object type errors]

**Are you using the right _input_ that the function expects?**

--

```{r, echo = FALSE}
decorate("sqrt('1')", eval = FALSE) %>%
  flair("'1'")
```

```{r, echo = FALSE, error = TRUE}

sqrt('1')

```

---

class: center, middle

.larger[Object type errors]

**Are you expecting the right _output_ of the function?**

--

```{r closure, echo = FALSE, error = TRUE, eval = FALSE}

my_obj <- seq(from = 1, to = 10, by = 1)

my_obj(5)

```

```{r, echo = FALSE}
decorate("closure", error = TRUE) %>%
  flair("my_obj(5)")
```

---

class: center, middle, inverse

.huge[Errors, warnings, and messages]

---

class: center, middle 

.larger[Messages]

--

Just because you see scary red text, this does **not** mean something went
wrong! This is just **R** communicating with you.

--

For example, you often see:

```{r, message = TRUE}
library(tidyverse)
```

---

class: center, middle

.larger[Warnings]

--

Often, **R** will give you a **warning**.

--

This means that your code *did* run...

--

...but you probably want to make sure it succeeded.

---

class: center, middle

.larger[Does this look right?]


```{r}
my_vec <- c("a", "b", "c")

my_new_vec <- as.integer(my_vec)
```

--

```{r}
my_new_vec
```

---

class: center, middle

.larger[Errors]

If the word **Error** appears in your message from **R**, then you have a
problem.

--

This means your code **could not run**!

--

```{r, error = TRUE}
my_vec <- c("a", "b", "c")

my_new_vec <- my_vec + 1
```

---

class: center, middle, inverse

.huge[Parlez-vous ERROR?]

---

.larger[**R** says...]

> Error: Object `some_obj` not found.

It *probably* means...

> You haven't run the code to create `some_obj`  OR you have a typo in the name!

--

```{r, error = TRUE}
some_ojb <- 1:10

mean(some_obj)
```

---

.larger[**R** says...]

> Error: Object of type 'closure' is not subsettable.

It *probably* means...

> Oops, you tried to use square brackets on a function

--

```{r, error = TRUE}
mean[1, 2]
```

---

.larger[**R** says...]

> Error: Non-numeric argument to binary operator

It *probably* means...

> You tried to do math on data that isn't numeric.

--

```{r, error = TRUE}
"a" + 2
```

---

.larger[What if none of these solved my error?]

--

Look at the help file for the function!

--

When all else fails, **Google** your error message.

--

- Leave out the specifics.

- Include the function you are using.

---

class: center, middle 

.larger[Try it...] 

**What's wrong here?**

</br>

```{r, error = TRUE}
matrix(c("a", "b", "c", "d"), num_row = 2)
```

