---
title: "Reshaping and Joining Data Frames"
author: "Week 4"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
```


In this lesson, you will learn:

* The importance of nicely formatted code and how to adhere to 
style "guidelines"

* How to **reshape** your dataset by pivoting *longer* or *wider*

* How to use **mutating joins** and **filtering joins** to combine different
datasets with related information.


---

`r emo::ji("play button")` Watch Videos: 25 minutes  

`r emo::ji("book")` Readings: 30-60 minutes   

`r emo::ji("computer")` Activities: 45-60 minutes  
  
`r emo::ji("check")` Check-ins: 2


---

## Code Style 

---

Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread.

In this section, you will explore the *tidyverse Style Guide*, an example of a
guideline for how you should format R code. Every style guide is opinionated. 
Some guidelines genuinely do make code easier to read, while others may seem 
more arbitrary. The most important feature about style guides is consistency. 
This consistency benefits you, since you have to make fewer decisions, and
also the people reading your code! 

The document is lengthy (since it covers **any** code you could write in R), so
we will focus on the sections that address aspects of coding in R we have 
already learned. We will use other sections as we build more skills!

[The tidyverse style guide -- Syntax](https://style.tidyverse.org/syntax.html) 

- __Focus on:__ Object names, Spacing, Function calls, Long lines, Assignment, 
Data, and Comments

[The tidyverse style guide -- Pipes](https://style.tidyverse.org/pipes.html)

- This is all important! 

[The tidyverse style guide -- ggplot2](https://style.tidyverse.org/ggplot2.html)

- This is all important! 

---

`r emo::ji("check")` **Check-in: _Code Style_**

**Question 1:** Based on the tidyverse style guide, where should whitespace
(blank spaces) go? 

- Before and after every `+` sign

- Before and after every `%>%`

- Before and after every athematic operator (`+`, `-`, `*`, `/`)

- After every comma

- Before every comma

**Question 2:** How many characters per line should you strive to have in your 
R code (and RMarkdown files)?

**Question 3:** Based on the styleguide, which of the following are **bad**
object names?

- `data`

- `cereals_data`

- `cereals.data`

- `cerealsData`

- `cerealsdata`

- `cereals_data1`

- `cereals_data_clean`

---

## Reshaping Data

---

### Tidy Data and Reshaping

`r emo::ji("book")` [Required Reading: *R4DS* 12.1-12.3 (Tidy Data)](https://r4ds.had.co.nz/tidy-data.html)

`r emo::ji("play button")` [Required Video: Data Layouts -- 7 minutes](http://somup.com/crfFrmF6VQ)

`r emo::ji("play button")` [Required Video: Pivoting Longer -- 9 minutes](http://somup.com/crfF3NF6hg)

`r emo::ji("play button")` [Required Video: Pivoting Wider -- 9 minutes](http://somup.com/crfqebF63k)

`r emo::ji("book")` [Optional Reading: *R4DS* 12.4-12.5 (Separate/Unite and Missing Values)](https://r4ds.had.co.nz/tidy-data.html) -- A few more tricks for data cleaning / wrangling, if you're interested!


---

`r emo::ji("check")` **Check-in: _Pivoting_**

- Download the `cereal.csv` data here: [https://app.box.com/s/zcr1fijnwxs27nvtxn4p38p3sgtxl4bt](https://app.box.com/s/zcr1fijnwxs27nvtxn4p38p3sgtxl4bt)

**Question 1:** Create a new dataset called `cereals_long`, that has three
columns:

* The `name` of the cereal

* A column called `Nutrient` with values `"protein"`, `"fat"`, or `"fiber"`.

* A column called `Amount` with the corresponding amount of the nutrient.

```{r, include = FALSE}

cereals <- read_csv(here::here("4_tidyr-joins", "Coursework", "Slides", 
                               "cereal.csv")
                    )

cereals_long <- cereals %>% 
  select(name, protein:fiber) %>% 
  pivot_longer(protein:fiber, 
               names_to = "Nutrient", 
               values_to = "Amount")
  
```


---

## Joining data

---

`r emo::ji("book")` [Required Reading: *R4DS* Chapter 13 (Relational Data)](https://r4ds.had.co.nz/relational-data.html)


`r emo::ji("book")` [Required Reading: Mutating Joins](https://www.garrickadenbuie.com/project/tidyexplain/)

  * These notes have great pictures / gifs for what is happening with joins! 
  * Note: We **will not** be using `spread()` and `gather()` in this course, as
  they are depricated functions!

`r emo::ji("warning")` **Disclaimer:** Because these chapters do a __much__
better job visualizing the concepts, I've chosen not to record additional
videos. 


---

`r emo::ji("check")` **Check-in: _Practice with Joins_**

The following code creates three datasets:

```{r}
prof_info <- data.frame(
  professor = 
    c("Bodwin", "Glanz", "Carlton", "Sun", "Theobold"),
  undergrad_school = 
    c("Harvard", "Cal Poly", "Berkeley", "Harvard", "Colorado Mesa University"),
  grad_school = 
    c("UNC", "Boston University", "UCLA", "Stanford", "Montana State University")
)

prof_course <- data.frame(
  professor = c("Bodwin", "Glanz", "Carlton", "Theobold"),
  Stat_331 = c(TRUE, TRUE, TRUE, TRUE),
  Stat_330 = c(FALSE, TRUE, TRUE, FALSE),
  Stat_431 = c(TRUE, TRUE, FALSE, TRUE)
)

course_info <- data.frame(
  course = c("Stat_331", "Stat_330", "Stat_431"),
  num_sections = c(8, 3, 1)
)

```

Here is what they look like once created:

```{r, echo = FALSE}
prof_info

prof_course

course_info
```

These datasets contain information about five Cal Poly professors, their
educational history, the classes they are able to teach, and the number of
sections of each class that need to be assigned.

**a) Combine datasets 1 and 2 to make this:**

```{r, echo = FALSE, message = FALSE}

prof_info %>% 
  inner_join(prof_course)

```

```{r, include = FALSE}
## OR 

prof_info %>%
  right_join(prof_course)

```

**b) Combine datasets 1 and 2 to make this:**

```{r, echo = FALSE, message = FALSE}

prof_info %>% 
  left_join(prof_course, by = "professor")

```

```{r, include = FALSE}
## OR 

prof_info %>% 
  full_join(prof_course, by = "professor")

```


**c) Combine datasets 2 and 3 to make this:**

```{r, echo = FALSE, message = FALSE}
prof_course %>%
  pivot_longer(Stat_331:Stat_431, 
               names_to = "course",
               values_to = "can_teach") %>%
  inner_join(course_info)

```

```{r, include = FALSE}
## OR

prof_course %>%
  pivot_longer(Stat_331:Stat_431, 
               names_to = "course",
               values_to = "can_teach") %>%
  right_join(course_info)

## OR

prof_course %>%
  pivot_longer(Stat_331:Stat_431, 
               names_to = "course",
               values_to = "can_teach") %>%
  left_join(course_info)

```

