---
title: "Learning More About Object Structures"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(xaringan)
library(showtext)
library(fontawesome)
library(tidyverse)
library(flair)

style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

.larger[Object Structures]

Is your object a...

* `vector`?

* `list`?

* `data.frame`?

* `tibble`?

* `matrix`?

* `function`? 

---

.larger[Vectors]

A **vector** is a set of values that are all the same **type**

```{r, error  = TRUE}
my_vec <- c("a", "b", "c")

str(my_vec)

my_vec <- c(1, 3, 5)

str(my_vec)
```

---

.larger[Atomic Vectors]

A vector with only **one** element is called an *_atomic vector_.

```{r, error  = TRUE}

my_vec <- 1

```

```{r}
str(my_vec)

my_vec[1]

```

---

.larger[Matrices]

A **matrix** is a **two-dimensional** set of vectors. 

- Every vector must have the same data type.

```{r, error  = TRUE}
my_mat <- matrix(c("a", "b", "c", "d"), nrow = 2, ncol = 2)

str(my_mat)
```

--

Note: If turning  vector into a matrix, be careful with rows and columns!

```{r}
my_mat
```


---

.larger[Lists]

A **list** is the the **most general** object.  

Any structure and any data type can be stored in a list!

```{r}
my_list <- list(
  a = "Well hi there",
  b = 1:10,
  c = my_mat
)
```

---

.bitlarger[Lists]

```{r}
my_list

str(my_list)

```

---

.bitlarger[Accessing Elements (Easy)]

```{r}
my_list$a
```

---

.bitlarger[Accessing Elements (Hard)]


```{r}
my_list[[1]]

my_list[1]
```

</br>
<center>

**What is the difference?**

---

.bitlarger[Accessing Elements (Hard)]

```{r}
str(my_list[[1]])
```

--

```{r}

str(my_list[1])
```