---
title: "Factors, Strings, and Dates"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("couRsework-img_path" = here::here("Images", "icons")) 
### change this if icons local
```

```{r, include = FALSE}
library(tidyverse)
library(lubridate)
```


In this lesson, you will learn:

* The importance of nicely formatted code and how to adhere to 
style "guidelines"

* How to **reshape** your dataset by pivoting *longer* or *wider*

* How to use **mutating joins** and **filtering joins** to combine different
datasets with related information.


---

`r emo::ji("play button")` Watch Videos: 60 minutes  
  * `r emo::ji("link")` Link to **all** Slides: [here])()

`r emo::ji("book")` Readings: 45-60 minutes   

`r emo::ji("computer")` Activities: 75-120 minutes  
  
`r emo::ji("check")` Check-ins: 6


---

# Factors

`r emo::ji("book")` [Required Video: Factors with forcats]()

`r emo::ji("book")` [Required Reading: *R4DS* -- Factors with forcats](https://r4ds.had.co.nz/factors.html)

---

`r emo::ji("book")` **Check-in: Functions from forcats 

Answer the questions based on the reading, video, **forcats**
[website](https://forcats.tidyverse.org/reference/index.html), or 
**forcats** [cheat sheet](https://www.rstudio.com/resources/cheatsheets/). 

* What is the difference between `fct_relevel()` and `fct_recode()`?

* What is the difference between `fct_collapse()` and `fct_lump()`?

* What is the difference between `fct_reorder()` and `fct_inorder()`?

*Feel free to share answers with your classmates in the Discord server, to make
sure you understand.*

`r emo::ji("check")` **Check-in: Apply your skills 

In your *colleges_clean* dataset, adjust the `REGION` variable to have nicer or
clearer labels and/or better categories.

---

# Strings

`r emo::ji("play")` [Required Video -- Strings with stringr]() 

`r emo::ji("book")` [Required Reading: *R4DS* -- Strings]"https://r4ds.had.co.nz/strings.html)

---

`r emo::ji("check")` **Check-in**: Functions from stringr

Answer the questions based on the reading, video, **stringr**
[website](https://stringr.tidyverse.org/reference/index.html), or 
**stringr** [cheat sheet](https://www.rstudio.com/resources/cheatsheets/). 

* What is the difference between `length()` and `str_length()`?

* What is the difference between `str_replace()` and `str_replace_all()`?

* What is the difference between `str_trim()` and `str_trunc()`?

* What is the difference between `str_sub()` and `str_subset()`?

* What does the `collapse` argument do in `str_c()`?

*Feel free to share answers with your classmates in the Discord server, to make
sure you understand.*

`r emo::ji("check")` **Check-in**: Apply your skills

Using your *colleges_clean* dataset, answer these questions:

**Question 1:** How *exactly* does Cal Poly's name appear in this dataset?

**Question 2:** How many colleges have the word "University" in their name?

**Question 3:** Which college has the *longest* name?  The *shortest* name?

---

# Dates

`r emo::ji("play")` [Required Video: DateTime Objects with lubridate]()

`r emo::ji("book")` [Required Reading: *R4DS* -- Dates and Times](https://r4ds.had.co.nz/dates-and-times.html)

**Extra Resources**

* [A more in-depth discussion of the POSIXlt and POSIXct data classes.](https://www.cyclismo.org/tutorial/R/time.html)

* [A tutorial on lubridate](https://data.library.virginia.edu/working-with-dates-and-time-in-r-using-the-lubridate-package/) - scroll down for details on intervals if 
you have trouble with `%within%` and `%--%`

---

`r emo::ji("check")` **Check-in**: Functions from lubridate

Answer the questions based on the reading, video, **lubridate**
[website](https://lubridate.tidyverse.org/reference/index.html), or 
**lubridate** [cheat sheet](https://www.rstudio.com/resources/cheatsheets/). 

- What is the difference between `year()` and `years()`?

<!-- year() -- extracts the year of a datetime object -->
<!-- years() -- creates a period object based on number of years  -->

- What `tz` would you use for San Luis Obispo? 

<!-- tz = "America/Los_Angeles" -->

- What is the difference between `%within%` and `%--%`?

<!-- %within% checks if a date is included in an interval -->
<!-- %--% creates an interval with a start and end time  -->
 
- What happens if you use the `parse_datetime()` function,
but you do not provide a `day` argument?

<!-- Assigns the first day of that month! -->


`r emo::ji("check")` **Check-in**: Apply your skills

(Answer the following using code (even if you know the answer already).

First, use the `make_date()` function to save your birthday as a
Datetime object in R.  

- What day of the week does/did your birthday fall on in 2021?

- On what day of the week should you plan your 50th birthday party?

- Find the date that is exactly 9 months before the day you were born.  If it falls on a holiday, think about the implications.
