---
title: "Practice Activity: Regression"
output: rmdformats::html_docco

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE, eval = FALSE)

library(tidyverse)
```

```{r, eval = FALSE}
dat <- read_csv(here::here("9_simulation-regression", 
                           "Practice Activity", 
                           "PA-regression", 
                           "data_2.csv"))

new_dat <- dat %>%
  select(-z) %>%
  mutate(
    y = y - mean(y),
    weight_before = x,
    weight_after = 1.2*x + y/5
  )

new_dat %>%
  ggplot(aes(x = weight_before, y = weight_after)) +
  geom_point()

mod <- lm(weight_after ~ weight_before, data = new_dat)

new_dat %>%
  modelr::add_predictions(mod) %>%
  ggplot(aes(x = weight_before, y = weight_after - pred)) +
  geom_point()

new_dat %>%
  select(weight_before, weight_after) %>%
  write_csv(here::here("8_Regression_Testing",
                       "Practice Activity", 
                       "PA-regression", 
                       "mystery_animal.csv"))
```

# Data

The dataset "mystery_animal.csv" contains the weights of a particular animal before and after a year of eating only roasted duck.


# Linear Regression 

Fit a linear regression to determine if the Duck Diet is associated with the 
animal gaining weight, losing weight, or neither.


# Residuals  

Plot the residuals (observed value minus predicted value) of your linear model versus `weight_before`. The `.resid` variable contains the residuals from the 
linear model you fit (assuming you stored it in an object!). 


# Mystery Animal

What animal do you see? 
