---
title: "Practice Activity 9 -- Regression"
author: "Solution by Dr. Theobold"
date: "February 28, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(kableExtra)
```

# Data 

The dataset "mystery_animal.csv" contains the weights of a particular animal before and after a year of eating only roasted duck.

```{r, include = FALSE}
animal <- read_csv(here::here("9_simulation-regression", 
                              "Practice Activity", 
                              "PA-regression", 
                              "mystery_animal.csv")
                   )
```


# Linear Model 

Compute a linear model to determine if the Duck Diet caused the animal to gain weight, lose weight, or neither.

```{r}
animal_lm <- animal %>% 
  lm(weight_after ~ weight_before, data = .)

summary(animal_lm) %>% 
  tidy() %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", 
                position = "center")
```

***The animal gained weight, since the coefficient associated with 
`weight_before` is positive.***


# Residuals

Plot the residuals (observed value minus predicted value) of your linear model 
versus `weight_before`. 

```{r}
animal_lm %>% 
  ggplot(aes(x = weight_before, y = .resid)) + 
  geom_point()

```

## That's definitely a shark.
