---
title: "Cereals Activity"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(flair)
library(palmerpenguins)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```


## In `Practice.Rmd`, make a new dataset called `kellogg_calories`, with the following changes:

1. There is a new variable called `cal_per_cup` which shows the number of calories in each cup of cereal.

2. The dataset only includes this new variable, the name of the cereal, and the brand of the cereal.

3. The dataset only includes Kellogg brand cereals

4. The cereals in the dataset are ordered from most to least calories per cup.

**Try to do all these in one sequence of pipes!**

5. Then choose one of the four main plot types (histogram, barplot, boxplot, scatterplot), make an interesting visualization from your `kellog_calories` dataset.

---
class:center,middle,inverse

# Grouping and summarizing
---
# `group_by`

* We have learned how to use the `summarize` functions to get information.

* What if you want to summarize your data within categories?

```{r}
penguins %>%
  group_by(species) %>%
  summarize(mean(body_mass_g))

```
---
# Warning!

## You should **always** have a `summarize` after a `group_by`!!!

(There are other ways to use it, but wait until you are very familiar with `dplyr` to try that.)

---
class: center, middle, inverse

# Your Turn!

---
# Grouping and summarizing

* Find the average calories per serving for each manufacturer.

* Find the average sugar per serving for hot versus cold cereals.

---

