---
title: "Introduction to ggplot"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
```

Now that we've refreshed how to work with different types of objects in R, let's
start visualizing!

> "The simple graph has brought more information to the data analysist's mind
> than any other device."
> -- John Tukey

`r emo::ji("play button")` Watch Videos: 50 minutes  

  - I've broken the videos up into smaller ~5 minute increments! 

`r emo::ji("book")` Readings: 45 minutes   

  - I've added an optional video on how to take notes in a programming textbook [here](http://somup.com/crffFPbHQ2). 

`r emo::ji("computer")` Activities: 30-45 minutes  
  
  - You'll be working through the textbook exercises for each section. 

`r emo::ji("check")` Check-ins: 1 (broken out by sections)


## Loading in the **ggplot2** Package

This week we will be taking a deep dive into data visualization, using the 
**ggplot2** package. The ggplot2 package is included in the **tidyverse**
package that you have already downloaded! To load in the ggplot2 package, you
will either need to run the code to load in the entire tidyverse package *or*
load in the ggplot2 package directly. 

```{r, eval = FALSE, echo = TRUE}
library(tidyverse)

OR 

library(ggplot2)
```

## Data Visualization with **ggplot2** 

This coursework will walk you through the different sections of Chapter Three 
in *R for Data Science*. 

`r emo::ji("book")` [Required Reading: Data Visualisation](https://r4ds.had.co.nz/data-visualisation.html). 


## Getting Started with **ggplot2** 

`r emo::ji("play button")` [Required Video: Why ggplot2? (4 minutes)](https://www.youtube.com/watch?v=5c-8yPCwVzs)

`r emo::ji("play button")` [Required Video: ggplot2 First Steps (4 minutes)](https://www.youtube.com/watch?v=1Jmbfh6GotI)

`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2 

**Question 1:** With the `mpg` dataset, make a scatterplot of `hwy` versus
`cty`. Which of these correctly describe the relationship seen in the plot? 

- positive
- strong
- negative
- weak
- linear
- non-linear

**Question 2:** What specifically does the code `ggplot(data = mpg)` do? 

- Creates a blank plot
- Makes a scatterplot of the first two variables in the `mpg` dataset
- Creates a ggplot with the `mpg` dataset

</br>

`r emo::ji("play button")` [Required Video: What is an aesthetic? (7 minutes)](https://www.youtube.com/watch?v=oYVuzW_t93w)

`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

**Question 3:** What aesthetics does this plot contain?

- x
- y
- color
- data


**Question 4:** Which of following changes would color the points blue? 

```
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, 
           color = blue))
```

```
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy) 
           color = "blue")
```

```
ggplot(data = mpg, mapping = aes(color = "blue)) +
geom_point(mapping = aes(x = displ, y = hwy))
```

</br>

`r emo::ji("play button")` [Required Video: What is a geometric object?](https://www.youtube.com/watch?v=pMC2l2YhjMI)


`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2

**Question 4:** Match each plot with the `geom_XXX()` function used to create
it! 

- Line Chart
- Boxplot 
- Histogram
- Area Chart

- `geom_point()`
- `geom_line()`
- `geom_histogram()`
- `geom_area()`

</br> 

`r emo::ji("computer")` [Required Tutorial: Data Visualization Basics](https://rstudio.cloud/learn/primers/1.1)

After working through the tutorial, if you feel more detail would be helpful,
consider the following readings:

* `r emo::ji("question")` [Modern Dive: Chapter 2](https://moderndive.com/2-viz.html)


## Getting a Bit Fancier

In this section we dig a bit deeper into how we can make our code more efficient
and other ways we can add additional variables to our plots. 

`r emo::ji("play button")` [Required Video: Global vs. local aesthetics (4 minutes)](http://youtu.be/y3BWlYa7q-Q?hd=1)

`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2 

**Question 5:** Match the code to the type of aesthetics that are being used:
Global or Local

```
ggplot(data = mpg, mapping = aes(x = mpg, y = hwy)) + geom_point()
```

```
ggplot(data = mpg) + geom_point(mapping = aes(x = mpg, y = hwy))
```

</br> 

`r emo::ji("play button")` [Required Video: Adding a third variable with facets](http://somup.com/crffbHbH1h)


## Data Transformations

`r emo::ji("play button")` [Required Video: What is being plotted? (6 minutes)](http://somup.com/crffb8bH1x)

`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2

**Question 6:** Match the stat function with its associated geom function. 

- `stat_count()`
- `stat_identity()`

- `geom_bar()`
- `geom_col()`

</br>

`r emo::ji("play button")` [Required Video: How do I change positions?](http://somup.com/crffFibH1R)

`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2

**Question 7:** What arguments for `geom_jitter()` control the amount of 
jittering? 

- width
- height
- stat
- position
- na.rm
- show.legend
- data
- mapping

</br>

## Plot Customizations 

`r emo::ji("play button")` [Required Video: Axis labels, color themes, and legends, oh my! (7 minutes)](http://somup.com/crffFDbHiU)

`r emo::ji("check")` Check-in: Introduction to Data Visualization with ggplot2

**Question 8:** What can the `labs()` function do? (select all that apply)

- Customize the x-axis label
- Customize the y-axis label
- Customize the legend label
- Customize the plot title
- Add a caption to the plot
