---
title: "Introduction to ggplot"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
```

Now that we've refreshed how to work with different types of objects in R, let's
start visualizing!

> "The simple graph has brought more information to the data analysist's mind than 
> any other device."
> -- John Tukey

`r emo::ji("play button")` Watch Videos: 50 minutes  

  - I've broken the videos up into smaller ~5 minute increments! 

`r emo::ji("book")` Readings: 45 minutes   

  - I've added an optional video on how to take notes in a programming textbook [here](http://somup.com/crffFPbHQ2). 

`r emo::ji("computer")` Activities: 30-45 minutes  
  
  - You'll be working through the textbook exercises for each section. 

`r emo::ji("check")` Check-ins: 1 (broken out by sections)


## Loading in the **ggplot2** Package

This week we will be taking a deep dive into data visualization, using the 
**ggplot2** package. The ggplot2 package is included in the **tidyverse** package
that you have already downloaded! To load in the ggplot2 package, you will 
either need to run the code to load in the entire tidyverse package *or* load in 
the ggplot2 package directly. 

```{r, eval = FALSE, echo = TRUE}
library(tidyverse)

OR 

library(ggplot2)
```

## Data Visualization with **ggplot2** 

This coursework will walk you through the different sections of Chapter Three 
in *R for Data Science*. 

`r emo::ji("book")` [Required Reading: Data visualisation](https://r4ds.had.co.nz/data-visualisation.html). 


## Getting Started with **ggplot2** 

`r emo::ji("play button")` [Required Video: Why ggplot2?](https://www.youtube.com/watch?v=5c-8yPCwVzs)

`r emo::ji("play button")` [Required Video: First Steps](https://www.youtube.com/watch?v=1Jmbfh6GotI)

`r emo::ji("check")` Check-in: 

**Question 1:** With the `mpg` dataset, make a scatterplot of `hwy` versus `cty`. 
Which of these correctly describe the relationship seen in the plot? 

**Question 2:** What specifically does the code `ggplot(data = mpg)` do? 

</br>

`r emo::ji("play button")` [Required Video: What is an aesthetic?](https://www.youtube.com/watch?v=oYVuzW_t93w)

`r emo::ji("check")` Check-in: 

**Question 3:** Exercise 3.3.1 #1: What's gone wrong with this code? Why are the
points not blue?

</br>

`r emo::ji("play button")` [Required Video: What is a geometric object?](https://www.youtube.com/watch?v=pMC2l2YhjMI)


`r emo::ji("check")` Check-in: 

**Question 4:** Match each plot with the `geom_XXX()` function used to create it! 

</br> 

`r emo::ji("computer")` [Required Tutorial: Data Visualization Basics](https://rstudio.cloud/learn/primers/1.1)

After working through the tutorial, if you feel more detail would be helpful,
consider the following readings:

* `r emo::ji("question")` [Modern Dive: Chapter 2](https://moderndive.com/2-viz.html)


## Getting a bit Fancier

In this section we dig a bit deeper into how we can make our code more efficient
and other ways we can add additional variables to our plots. 

`r emo::ji("play button")` [Required Video: Global vs. local aesthetics](http://youtu.be/y3BWlYa7q-Q?hd=1)

`r emo::ji("check")` Check-in: 

**Question 5:** Match the code to the type of aesthetics that are being used: global or local

</br> 

`r emo::ji("play button")` [Required Video: Adding a third variable with facets](http://somup.com/crffbHbH1h)

`r emo::ji("check")` Check-in: 

**Question 6:** What do the `nrow` and `ncol` arguments to `facet_wrap()` do?


## Data Transformations

`r emo::ji("play button")` [Required Video: What is being plotted?](http://somup.com/crffb8bH1x)

`r emo::ji("check")` Check-in: 

**Question 7:** Match the stat function with its associated geom function. 

</br>

`r emo::ji("play button")` [Required Video: How do I change positions?](http://somup.com/crffFibH1R)

`r emo::ji("check")` Check-in: 

**Question 8:** What parameters to `geom_jitter()` control the amount of 
jittering? 


## Plot Customizations 

`r emo::ji("play button")` [Required Video: Axis labels, color themes, and legends, oh my!](http://somup.com/crffFDbHiU)

`r emo::ji("check")` Check-in: 

**Question 9:** What can the `labs()` function do? (select all that apply)
