---
title: "Introduction to Reading in Data"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
```

`r emo::ji("play button")` Watch Videos: 14 minutes  

`r emo::ji("book")` Readings: 15-20 minutes   

`r emo::ji("computer")` Activities: 30 minutes  

`r emo::ji("check")` Check-ins: 1


## Loading Data Into R 

### Where does my data live?

`r emo::ji("play button")` [Required Video: Data Locations and the {here} package (8 minutes)](https://www.youtube.com/watch?v=jvWwrELLEw8)

### How do I load my data?

`r emo::ji("book")` [Required Reading: Data Import](https://r4ds.had.co.nz/data-import.html)

`r emo::ji("play button")` [Required Video: Loading Data (6 minutes)](https://www.youtube.com/watch?v=Kp4tJYCAvUs) 

`r emo::ji("computer")` Download the `ages.zip` folder and the 
`data_import_practice.qmd` files on Canvas. Save these files in
your Stat 331 folder, possibly inside a subfolder named "Activities" or
"Week 2". 

**Note:** You will need to **extract** the contents of the `ages.zip` file, 
that means you will need to uncompress the files from the folder for RStudio 
to know where to get the data from. 

Now, open your STAT 331 RStudio Project. Next, open the
`data_import_practice.qmd` file and write the code necessary to carry out the
following tasks:

1. Read in the dataset `ages.csv`.
2. Read in the dataset `ages_tab.txt`
3. Read in the dataset `ages_mystery.txt`
4. Read in the dataset `ages.xlsx`
5. Both `read.csv()` and `read.table()` have an optional argument called
`stringsAsFactors`.  Figure out what this does.
6. Find a way to use `read_csv()` to read `ages.csv` with the variable "Name" as
a factor and "Age" as a character.


---

`r emo::ji("check")` Check-in: Loading Data 

**Question 1**

What line of code did you use -- **including the `here()` function** -- to read
in `ages.csv` dataset?

**Question 2**

What does the `stringsAsFactors` argument in `read.csv()` do?
