---
title: "Lab 3 Grading Guide"
output: html_document
---

</br>

**If there is no response to a question a score of 0 is awarded.**

**If code outside of the _R for Data Science_ book is used, a citation must be 
included. If no citation is included, the question is awarded a 0.**

</br>

# Question 0

-	4: Reads data in with the `here()` function 
-	Redo: If uses relative or full path to read in the data

</br>

# Question 1

**4**: Must have the following,

  - number of participants (168)
  - context of participants (undergraduate students)
  - context of study (familiarity with African-American English)
  - description of how familiarity was defined (1 = not at all, 5 = well)

**3**: Has all of the above except for one

**Redo if above criteria is not met**

</br>

# Question 2

**4**: Rows are __one__ word for _each_ participant 
(168 participants $\times$ 64 words = 10,752 rows)

**Redo if above criteria is not met**

</br>

# Question 3

**4** 

  - Notes missing values for some variables were replaced with the mean
  - Benefits: fewer instances of missing data
  - Drawbacks: if the mean is not representative of what that individual would
  have responded

**3**: Meets the first two criteria, but does not discuss how the mean may
misrepresent the true value

**Redo if above criteria is not met**

</br>

# Question 4

**4**

- Uses `mutate()` to convert some variables (e.g. `sex`, `ethnic`) to factors
- Uses `filter()` with `is.na()` or `drop_na()` to specify variables to remove
missing values from 

**Redo if uses `drop_na()` without any variables specified** -- removing **all**
of the missing values

</br>

# Question 5

**4**: Use `distinct()` and `count()` to get 64 words

**3**: Use a “base R” method (e.g. `unique()`) to get 64 words

**Redo if above criteria is not met**

</br>

# Question 6

**4**

  - Use `mutate()` and `if_else()` to create a new variable based on values of
  `ethnic` variable  
  - Note: Can use `ifelse()` function, but note that the behavior of `if_else()`
  is preferred!

```{r, eval = FALSE}
hiphop <- hiphop %>%
  mutate(
    ethnic_group = if_else(ethnic == "white",
                           "white", 
                           "non-white"),
    ethnic_group = as_factor(ethnic_group)
  )
```

**Redo if above criteria is not met**

</br>

# Question 7

__This answer will depend on how they “cleaned” their data.__

**Every solution must use `distinct(subj, .keep_all = TRUE)` to select distinct
study subjects** 


**4**: Discusses all  
**3**: Discusses two  
**Redo**: Discusses one or none  

  - number of female / male participants
  - number of participants in ethnic groups 
  - summaries for the ages of the participants (e.g. min, max, mean, median)

**Note:** Looking at white / non-white groups created in #6 **does not** satisfy
the second summary of `ethnic` groups

</br>

# Question 8 

**4**: Produces at least two appropriate plots

**3**: Produces two appropriate plots, but forgets to change axis labels and / 
or plot titles

**Redo**: One of more of the plots produced are not appropriate

__List of Appropriate Plots__

-	Histogram of `age`s -- can add colors / facets for sex or ethnic variable(s) 

- Density Ridges of `age`s -- can add colors / facets for sex or ethnic
variable(s) 
  
-	Boxplot of `age`s: can add colors / facets for sex or ethnic variable(s)

-	Barplot of `sex` and / or ethnicity variables -- can add colors / facets for
other variable(s)

</br>

---

# Familiar Words

### Part (a)

**4**: 

  * Uses `filter()` to remove observations under 20
  * Groups by `word`
  * Uses `summarize()` to calculate the mean `familiarity` score 
  * Uses `slice_min()` to calculate the least familiar
  * Uses `slice_max()` to calculate the most familiar

**3**: Uses all of the above tools, except uses an alternative method to find
the min / max

_Answer should be_:

  * “off the hook” most familiar
  * “catch the vapors” least familiar
  
**Redo**: If group does not use these 5 tools

</br>

### Part (b)

**4**:

  * Uses `filter()` to retain “Female” and “non-white” participants
  * Groups by `word`
  * Uses `summarize()` to calculate the mean `familiarity` score
  * Uses `slice_min()` to calculate the least familiar
  * Uses `slice_max()` to calculate the most familiar

**Note:** If team was previously deducted points for using a function other than
`slice_XXX()` to find the minimum / maximum, do not deduct points again! 

_Answer should be_: 

  * "feel me" most familiar
  * "break someone out" "dukey rope" "plex" and "rollie" least familiar
  
**Redo**: If group does not use these 5 tools

</br>

### Part (c)

**4**:

  * Uses `filter()` to retain `“Male”` and `“white”` observations over 30
  * Groups by `word`
  * Uses summarize() to calculate the mean `familiarity` score

**Note:** If team was previously deducted points for using a function other than
`slice_XXX()` to find the minimum / maximum, do not deduct points again! 

_Answer should be_: 

  * “5-0” most familiar
  * 25 terms are the least familiar
  
**Redo**: If group does not use these 5 tools

</br>

---

# Study Subjects

**4**: 

  - Uses the following `filter()`s: 

```{r, eval = FALSE}
filter(sex == "Male", 
       age <= 23, 
       age >= 17,
       city < 60000,
       city > 10000)
```

  - Uses `slice_max()` to find the maximum of the `bieber` variable

_Answer should be:_ Justin Bieber is subject "p17."

**3**: Uses all of the above `filter()`s, but uses an alternative method to find
the min / max of the `bieber` variable

**Redo if above criteria is not met**
