---
title: "Recap of Data Visualization with ggplot2"
date: "April 6, 2021"
author: "Dr. Allison Theobold"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---


```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE, 
                      fig.align = "center")

library(palmerpenguins)
library(tidyverse)
library(ggridges)
library(gridExtra)
```

.pull-left[
```{r, echo = FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = species, fill = sex)) + 
  geom_density_ridges(alpha = 0.5)

```
]

.pull-right[

</br>
</br> 

.hand[What are the `aes`thetics in this plot? 

What `geom`etric object is being plotted?]
]
---

.bitlarger[Univariate (One Variable) Visualizations -- For Numerical Data]

- Histogram 

--

- Boxplot

--

- Density Plot

---

.larger[Histogram]

.pull-left[
```{r, message = FALSE, fig.height = 4, fig.width = 5, warning = FALSE, fig.align = 'center'}
ggplot(data = penguins, 
         aes(x = bill_length_mm)) + 
  geom_histogram()
```
]

--

</br>
</br>

.pull-right[Is this aesthetic *global* or *local*?]
---

.pull-left[
.larger[Pros]
- Easy to inspect
- Higher bars represent where data are relatively more common 
- Inspect shape of a distribution (skewed or symmetric) 
- Identify modes
]

.pull-right[
.larger[Cons]
- **Do not** plot raw data, plot summaries (counts) of the data! 
- Sensitive to `binwidth`
]
---

.larger[Boxplot]

.pull-left[
```{r, fig.width = 5, fig.height = 4}
ggplot(data = penguins, 
       aes(x = bill_length_mm)) + 
  geom_boxplot()
```
]

.pull-right[

- What calculations are necessary to create a boxplot? 

- What are strengths of a boxplot? 

- What are weaknesses of a boxplot?

]

---

.larger[Density Plot]

.pull-left[
```{r, fig.width = 6, fig.height = 4}
ggplot(data = penguins, 
         aes(x = bill_length_mm)) + 
  geom_density()
```
]

.pull-right[
- A smooth approximation to a variable's distribution 
- Plots density (as a proportion) on the y-axis

</br> 

- What are the strengths of a density plot?  
- What are the weaknesses of a density plot? 
]

---

.bitlarger[Bivariate (Two Variables) Visualizations -- For Numerical Data]


- Side-by-Side Boxplots 

--

- Side-by-Side Density Plots (Ridge Plots)

--

- Scatterplots 

--

- Faceted Histograms 

---

.bitlarger[Side-by-Side Boxplots]

.pull-left[
```{r, fig.width = 6, fig.height = 5, fig.align = "center"}
ggplot(data = penguins, 
         aes(y = bill_length_mm, x = species)) + 
  geom_boxplot()
```
]

.pull-right[
Can stack vertically (with categorical variable on `y`-axis) **or** horizontally (with categorical variable on `x`-axis)
]

---

.bitlarger[Faceted Histograms] 

```{r, fig.width = 8, fig.height = 5, fig.align = "center", echo = TRUE}
ggplot(data = penguins, 
       aes(x = bill_length_mm)) + 
  geom_histogram(binwidth = 3) + 
  facet_wrap(~ species)
```

---

.bitlarger[Scatterplots]

```{r, fig.width = 6, fig.height = 5, fig.align = 'center'}
ggplot(data = penguins, 
       aes(y = bill_length_mm, x = bill_depth_mm)) + 
  geom_point()
```

---

.bitlarger[Multivariate Plots]

There are two main methods for adding a third (or fourth) variable into a 
data visualization: 

- Colors: 
  * creates colors for every level of a categorical variable
  * creates a gradient for different values of a quantitative variable

--

- Facets: 
  * creates subplots for every level of a variable
  * labels each sub-plot with the value of the variable

---

.bitlarger[Colors in Scatterplots]

```{r, fig.width = 6, fig.height = 4, fig.align = 'center'}
ggplot(data = penguins, 
         aes(y = bill_length_mm, 
             x = bill_depth_mm, 
             color = species)) + 
  geom_point()
```

---

.bitlarger[Colors in Boxplots]

```{r, fig.width = 6, fig.height = 4, fig.align = 'center'}
ggplot(data = penguins, 
         aes(y = bill_length_mm, 
             x = species, 
             color = sex)) + 
  geom_boxplot()
```

---

.bitlarger[Facets in Scatterplots] 

```{r, fig.width = 8, fig.height = 4, fig.align = 'center'}
ggplot(data = penguins, 
         aes(y = bill_length_mm, 
             x = bill_depth_mm, 
             color = sex)) + 
  geom_point() + 
  facet_wrap(~ species)
```

---

.bitlarger[Facets in Boxplots]

```{r, fig.width = 8, fig.height = 4, fig.align = 'center'}
ggplot(data = penguins, 
         aes(x = bill_length_mm,
             y = species,
             fill = sex)) + 
  geom_boxplot(alpha = 0.3) + 
  facet_wrap(~ island)
```

---

## Your Turn!

