---
title: ".hand[What is an **aes**thetic?]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include = FALSE}
library(ggplot2)
library(flair)
```

.larger[An aesthetic is a visual property of the objects in your plot.]

</br>

.large[
- variables on axes 
- size, shape, or the color of points

</br>

>"You can convey information about your data by mapping the aesthetics in your
> plot to the variables in your dataset.

]

---

.larger[Axes]

<center>
```{r axes, eval = FALSE, echo = FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r, echo = FALSE}
decorate_chunk(chunk_name = "axes", eval = FALSE) %>% 
  flair(pattern = "x = displ", color = "tomato") %>% 
  flair(pattern = "y = hwy", color = "blue")
```

</center>
</br> 

.large[

- Maps **`disp`** to `r flair("x-axis", pattern = "x-axis", color = "tomato")`
- Maps **`hwy`** to `r flair("y-axis", pattern = "y-axis", color = "blue")`
]

---

.larger[Colors]

.pull-left[
```{r color, echo = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ,
                  y = hwy,
                  color = class))
```
]

.pull-right[

```{r, echo = FALSE}
decorate_chunk(chunk_name = "color", eval = FALSE) %>% 
  flair(pattern = "color = class", color = "green")
```

- `r flair("colors", pattern = "color", color = "green")` points based on **`class`** of vehicles
]

---

.larger[Shape]

.pull-left[
```{r shape, echo = FALSE, warning = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ,
                  y = hwy,
                  shape = class))
```
]

.pull-right[
```{r, echo = FALSE}
decorate_chunk(chunk_name = "shape", eval = FALSE) %>% 
  flair(pattern = "shape = class", color = "purple")
```

- Uses a different `r flair("shape", pattern = "shape", color = "purple")` points based on **`class`** of vehicles
]

---

.larger[Size]

.pull-left[
```{r size, echo = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ, 
                  y = hwy, 
                  size = cyl))
```
]

.pull-right[
```{r, echo = FALSE}
decorate_chunk(chunk_name = "size", eval = FALSE) %>% 
  flair(pattern = "size = cyl", color = "red")
```

- Uses a different `r flair("size", pattern = "size", color = "red")` points based on **`ncyl`** of the vehicle
]

---

.larger["Plain" Aesthetics]

.pull-left[
.large[
Assign a single `color`, `shape`, or `size` to **all** points 
]
```{r, eval = FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), 
             color = "blue", 
             shape = "triangle",
             size = 3
             )
```
]

.pull-right[

```{r, echo = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), 
             color = "blue", 
             shape = "triangle",
             size = 3
             )
```
]