---
title: ".hand[How do I make my plot look nicer?]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include = FALSE}
library(tidyverse)
library(gridExtra)
library(flair)
library(viridis)
```

class: inverse, center

.larger[**The small things matter just as much!**] 

.bitlarger[
axis labels  
legend position  
color choice

]

---

.larger[Clearer labels with `labs()`]

.pull-left[
```{r labs, eval = FALSE, echo = FALSE}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, 
                            y = hwy, 
                            color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of Cylinders")

```

```{r, echo = FALSE}
decorate_chunk("labs", eval = FALSE) %>% 
  flair(pattern = "labs")
```

]

.pull-right[
```{r, echo = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, y = hwy, color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of \nCylinders")

```
]

---

.larger[Change the position of the legend with `theme()`]

.pull-left[
```{r theme, eval = FALSE, echo = FALSE}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, 
                            y = hwy, 
                            color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of Cylinders") + 
  theme(legend.position = "top")

```

```{r, echo = FALSE}
decorate_chunk("theme", eval = FALSE) %>% 
  flair(pattern = "legend.position")
```


]

.pull-right[
```{r, echo = FALSE, fig.width = 5, fig.height = 5}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, y = hwy, color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of \nCylinders") +
  theme(legend.position = "top") 
```
]

---

.larger[Change the colors with different packages]

.pull-left[Options:

- **RColorBrewer** 

- **viridis** 

- **ggsci**

- **wesanderson**
]

.pull-right[
```{r, echo = FALSE, fig.height = 5.5, fig.width = 5.5}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, y = hwy, color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of \nCylinders") +
  theme(legend.position = "top") + 
  scale_colour_viridis_c(option = "E")
```
]

