---
title: ".hand[What is a **geom**?]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include = FALSE}
library(tidyverse)
library(gridExtra)
library(flair)
```

.larger[How to represent the data?]

```{r, echo = FALSE, fig.width = 10, fig.height = 6, message = FALSE, fig.align='center'}
p1 <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) 
             
p2 <- ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
      
grid.arrange(p1, p2, nrow = 1)       


```

---

.larger[**geom**]

> "A **geom** is the geometrical object that a plot uses to represent data."   

> "People often describe plots by the type of geom that the plot uses."

</br>

.large[

| Plot                | geom            |
|---------------------|-----------------|
| Boxplot     | boxplot   |
| Bar plot    | bar       |
| Line graph  | line      |
| Scatterplot | point     |

]

---

.larger[Changing **geom**s]

```{r point, eval = FALSE, echo = FALSE}
# Scatterplot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

```

```{r, echo = FALSE}
decorate_chunk(chunk_name = "point", eval = FALSE) %>% 
  flair(pattern = "geom_point", color = "blue")
```

</br>

```{r smooth, eval = FALSE, echo = FALSE}
# Smoothing Curve
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r, echo = FALSE}
decorate_chunk(chunk_name = "smooth", eval = FALSE) %>% 
  flair(pattern = "geom_smooth", color = "green")
```
---

.larger[**geom**s and **mapping**s]

.large[Different `geom`s take different `mapping` arguments

.pull-left[
- `geom_smooth`: no `shape`, `linetype` instead

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 4}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = class))
```
]

.pull-right[

- `geom_bar`: `color` bad, `fill` better

```{r, echo = FALSE, fig.width = 6, fig.height = 4}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = drv, fill = class))
```
]
]

---

.larger[Multiple **geom**s in one plot!]

```{r, fig.align='center', message = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

