---
title: "Directories & Scientific Reproducibility"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(
  primary_color = "#26116c",
  secondary_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

---

class: center, middle

# Location of Data


---

# read.csv()

- The most common way to read data is with `read.csv()`.

- You may supply either a **url** to the dataset, or a **path**.

```{r read.csv, eval = FALSE}
hiphop <- read.csv("https://www.dropbox.com/s/1qqyshx5ikt9zoc/hiphop.csv?dl=1")
## This will work for everyone!

```

--

```{r read.csv2, eval = FALSE}

hiphop <- read.csv("/Users/kbodwin/Dropbox/Teaching/STAT-331-Online/Data/hiphop.csv")
## This will work only on my computer

```

--

```{r read.csv3, eval = FALSE}

hiphop <- read.csv("hiphop.csv")
## This will work if the file is in the same directory as the code (i.e., the project directory)
```

--

```{r read.csv4, eval = FALSE}

hiphop <- read.csv("./Data/hiphop.csv")
## This will work if a folder called "Data" is in the same directory as the code.
```
---
# The **here** package

A great solution to consistency in paths is the `here` package:

```{r, eval = FALSE}
install.packages("here")
```

--

This package thinks that *"here"* means *"the directory of the closest project"*.

```{r}
here::dr_here()
```
---
# The **here** package

Recall that these slides live in a subfolder for Week 2 and for this Lesson:

```{r}
getwd()
```

--
If I wanted to access a dataset from my main Data folder, I'd have to use a relative file path:

```{r, eval = FALSE}
hiphop <- read.csv("../../Data/hiphop.csv")
```
--

If I ever reshuffle my folder structure, I'm in trouble!

---
# The **here** package

Instead of trying to reference data relative to **this** file, let's refference it relative to the **Project** location, which is unlikely to change.


```{r}
here::here("Data", "hiphop.csv")
```

--


```{r, eval = FALSE}
hiphop <- read.csv(here::here("Data", "hiphop.csv"))
```

This will **adapt** to each person's path!

---
class: center, middle

# Reading in data

---

# Common types of data files

- `.csv` :  "Comma-separated"
```
Name, Age
Bob, 49
Joe, 40
```
- `.txt`:  Plain text
    + Could be just text
    + Could be comma-separated data
    + Could be tab-separated, bar-separated, etc.
    + Need to let R know what to look for

- `.xls, .xlsx`:  Microsoft Excel Spreadsheet
    + Common approach: save as `.csv`
    + Nicer approach:  `readxl` package


---
# read.csv() and read.table()

- A more general version of `read.csv()` is `read.table()`.

- Tell R what symbol separates data observations with `sep = `

```{r read.table, eval = FALSE}
hiphop <- read.table("hiphop.csv", sep = ",")
hiphop <- read.table("hiphop.csv", sep = "\t")  
## if tabs between data
```

---
# readr

The tidyverse has some cleaned-up versions in the `readr` and `readxl` packages: 

- `read_csv` works like `read.csv`, with some extra stuff

- `read_tsv` is for tab-separated data

- `read_table` is for any data with "columns" (white space separating)

- `read_delim` is for special "delimiters" separating data

- `read_excel` is specifically for dealing with Excel files

---
# Exercises

The folder "Age_Datasets" on Canvas several datasets containing the names and ages of five men.  The datasets are stored as different file types.  Save these in a reasonable place, then do the following:

1. Read in the dataset `ages.csv`.
2. Read in the dataset `ages_tab.txt`
3. Read in the dataset `ages_mystery.txt`
4. Read in the dataset `ages.xlsx`
5. Both `read.csv` and `read.table` have an optional argument called `stringsAsFactors`.  Figure out what this does.
6. Find a way to use `read_csv` to read `ages.csv` with the variable "Name" as a factor and "Age" as a character.
