---
title: "How do I read in my data?"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```


class: middle

.larger[Common types of data files]

.pull-left[

`.csv` :  "Comma-separated"
```
Name, Age
Bob, 49
Joe, 40
```
</br>

`.xls, .xlsx`:  Microsoft Excel Spreadsheet
  - Common approach: save as `.csv`
  - Nicer approach:  `readxl` package

]

.pull-right[
`.txt`:  Plain text
  - Could be just text
  - Could be comma-separated data
  - Could be tab-separated, bar-separated, etc.
  - Need to let R know what to look for
]

---

.larger[`read.csv()` and `read.table()`]

- A more general version of `read.csv()` is `read.table()`.

- Tell R what symbol separates data observations with `sep = `

--

```{r read.table, eval = FALSE}
ages <- read.table("ages.txt", sep = ",")
## comma delimiter 

ages <- read.table("ages.txt", sep = "\t")  
## tab delimiter
```

---

.larger[A Cleaner Approach with readr]

The tidyverse has some cleaned-up versions in the `readr` and `readxl` packages: 

- `read_csv()` works like `read.csv`, with some extra stuff

- `read_tsv()` is for tab-separated data

- `read_table()` is for any data with "columns" (white space separating)

- `read_delim()` is for special "delimiters" separating data

- `read_excel()` is specifically for dealing with Excel files

---

# Exercises

The folder "Age_Datasets" on Canvas several datasets containing the names and
ages of five individuals.  The datasets are stored as different file types.
Save these in a reasonable place, then do the following:

1. Read in the dataset `ages.csv`.
2. Read in the dataset `ages_tab.txt`
3. Read in the dataset `ages_mystery.txt`
4. Read in the dataset `ages.xlsx`
5. Both `read.csv` and `read.table` have an optional argument called `stringsAsFactors`.  Figure out what this does.
6. Find a way to use `read_csv` to read `ages.csv` with the variable "Name" as a factor and "Age" as a character.
