---
title: "An Introduction to the tidyverse & dplyr"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(xaringan)
library(showtext)
library(fontawesome)
library(tidyverse)
library(flair)
library(palmerpenguins)

style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

class: center, middle

# R History lesson:
# What is the `tidyverse`?

---
# The tidyverse

The `tidyverse` is a collection of packages that follow a (somewhat) consistent philosophy and structure.

The term `base R` is sometimes used as an alternative, to refer to only R's built-in packages.

---

# A brief history

--

**2008:** `plyr` and `ggplot` released as part of **Hadley Wickham**'s doctoral dissertation.

--

**2012:** The pipe (`%>%`) is created.  

*(While the pipe itself is not technically tidyverse, the tidyverse is built around its use.)*

--

**2014:** `dplyr` released; Wickham hired as Chief Scientist at **RStudio**

--

**Today:** A paid team of 20-ish developers maintains a huge number of packages.  (https://www.tidyverse.org/)

---

</br>
</br>

>Programs must be written for people to read, and only incidentally for 
> machines to execute.
>
>— Hal Abelson

</br>
</br>

```{r, eval = FALSE}
penguins %>%
  filter(species == "Adelie")
```

```{r, eval = FALSE}

penguins[penguins$species == "Adelie",]
```

---


# The Philosophy of the tidyverse

- Keep functions as simple as possible 
  * Each function should do one thing well 
  
--

- Function names should be verbs
  * When many functions use the same verb, focus on noun (ggplot2)
  * Favor explicit, lengthy names, over short, implicit, names

--

- Build **pipe-friendly** functions

---

# The tidyverse

**Pros:**

* Fairly consistent behavior of functions

    + Predictable **input** and **output** types
    + *"opinionated"*
    + More on this later
    
--

* Semi-guaranteed maintenance of packages
(because RStudio pays developers!)

--

* More available and polished documentation and tutorials

--

* Welcoming community

---
# The tidyverse

**Cons:**

* Sometimes less flexible for complex tasks

--

* Generally slower, especially on big data

--

* Updates frequently -> sometimes this makes your old code unrunnable.

--

* Some people object to **RStudio**'s strong influence over an open-source community.

---
class: center, middle

# Should you use the tidyverse???

--

Even the most die-hard tidyverse fanatics use base R and other packages quite frequently.

--

Above all, the best code approach is the one you enjoy the most and that gets the job done!

--

You are free to use whatever package(s) you want in this class!

--

However, most of my lessons and examples will be in the **tidyverse**.
