---
title: "Reproducibility: Directories and Deliverables"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
```

The theme of this lesson is good management of your files and data.

You will learn how to:

* Identify **folders** and **paths**.

* Create and use **R Projects**.

* **Load data** into R from files.

* Deal with package challenges.

* Create R Markdown documents with good **reproducible principles**.

---

`r emo::ji("play button")` Watch Videos: 25 min

`r emo::ji("book")` Readings: 15 minutes 

`r emo::ji("computer")` Activities: 10 min 

`r emo::ji("check")` Check-ins: 2

---

## Principles of Reproducibility

`r emo::ji("play button")` [Required Video: Directories, Paths, Projects (17 minutes)](http://somup.com/creu0fbzyn)

`r emo::ji("book")` [Required Reading: Workflow and Projects](https://r4ds.had.co.nz/workflow-projects.html) 

---

## Reproducible Deliverables

### Introduction to R Markdown

`r emo::ji("book")` [Required Reading: Intro to R Markdown](https://r4ds.had.co.nz/r-markdown.html)

`r emo::ji("play button")` [Required Video: Introduction to R Markdown by Dr. Kelly Bodwin (8 minutes)](https://www.youtube.com/embed/Zt6MJ2m2bNk)

<!-- UPDATE VIDEO WITH NEW SLIDES -->

`r emo::ji("play button")` [Additional Video: Design Choices for R Mardown Documents (11 minutes)](https://www.youtube.com/watch?v=w_HV_RI1lN8)

---

# `r emo::ji("question")` Extra Resources

* [Software Carpentry -- Project Managment with RStudio](http://swcarpentry.github.io/r-novice-gapminder/02-project-intro/index.html)

---

`r emo::ji("check")` Check-in: RStudio Projects

Take a screenshot of your class directory, showing the **R Project** and the
folder organization, including a **Week 1** or **Practice Activity** folder.

---

# `r emo::ji("question")` Extra Resources

* [Click Here](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
to download a cheatsheet for R Markdown.

* [This website](https://holtzy.github.io/Pimp-my-rmd/) has some tips for making
your R Markdown documents more beautiful.
