---
title: "Practice Activity 4: Tidy Data with tidyr"
author: "Stat 331"
output: rmdformats::readthedown
---

Today you will be tidying untidy data to explore the relationship between 
countries of the world and military spending. 

## Data for Today's Activity  

The SIPRI (Stockholm International Peace Research Institute) Military Expenditure
Database is an open source dataset with contains consistent time series on the 
military spending of countries for the period 1949–2018. The database is updated
annually, which may include updates to data for any of the years included in the
database.

Military expenditure in local currency at current prices is presented according
to both the financial year of each country and according to calendar year, 
calculated on the assumption that, where financial years do not correspond to 
calendar years, spending is distributed evenly through the year. 
Figures in constant (2017) and current US $, as a share of GDP and per capita 
are presented according to calendar year. 

The availability of data varies considerably by country, but for a majority of 
countries that were independent at the time, data is available from at least the
late 1950s. Estimates for regional military expenditure have been extended 
backwards depending on availability of data for countries in the region, but no 
estimates for total world military expenditure are available before 1988 due to 
the lack of data for the Soviet Union.

SIPRI military expenditure data is based on open sources only.


## Data Inspection

__Download the data [here](). Open the Excel file and inspect the file.__ 

First, you should notice that there are ten different sheets included in the 
dataset. We are interested in the sheet labeled "Share of Govt. spending" which 
contains information about the share of government spending that is allocated to
military spending. 

Next, you'll notice that there are notes about the dataset in the first seven
rows. Ugh! 

Rather than copying this one sheet into a new Excel file and deleting the first 
seven rows, let's learn something new about the `read_xlsx()` function! 


## Data Import

The `read_xlsx()` function has a `sheet` argument, where you specify the name of
the sheet that you want to use.  

**Note:** The name must be passed in as a string (in quotations)!

The `read_xlsx()` function also has a `skip` argument, where you specify the 
number of rows you want to be skipped *before* reading in the data.

__1.__ Modify the code below to read the military expenditures data into your
work space. 

```{r, message = FALSE}
library(readxl) 
library(tidyverse)

military <- read_xlsx(here::here("4-tidyr-joins", "Practice Activity", 
                                 "data", "gov_spending_per_capita.xlsx"), 
                      sheet = "Share of Govt. spending", 
                      skip = 7)
## This is MY path to the data! Yours will look different! 
```

## Data Cleaning 

In addition to `NA`s, missing or unavailable values were coded two ways.

__2.__ Find these two methods and write the code to replace these values with
NAs. Save the mutated dataset into a new object named `military_clean`.  


```{r, warning = FALSE}
military_clean <- military %>% 
                    mutate_at(vars(`1988`:`2019`),
                              ~na_if(.x, y = ". .")
                              ) %>%
                      mutate_at(vars(`1988`:`2019`),
                              ~na_if(.x, y = "xxx")
                              )
```


Because of the use of characters to mark missing values, all of the 
columns 1988 through 2019 were read in a characters. 

__3.__ Mutate these columns to all be numeric data types, instead of a character
data type. Save these changes into an updated version of `military_clean`.  


```{r}
military_clean <- military_clean %>% 
  mutate_at(vars(`1988`:`2019`), as.numeric) 
```

If you give the `Country` column a look, you'll see that there are names
of continents and regions included. These names are only included to make it 
simpler to find countries, as they contain no data. 

Luckily for us, these names were stored in the "Regional totals" sheet. We can 
use the `Region` column of this dataset to filter out the names we don't want. 

Run the code below to read in the "Regional totals" dataset, making any 
necessary modifications to the path. 

```{r, message = FALSE}
cont_region <- read_xlsx(here::here("4-tidyr-joins", "Practice Activity", 
                                 "data", "gov_spending_per_capita.xlsx"), 
                      sheet = "Regional totals", 
                      skip = 14) %>% 
  filter(Region != "World total (including Iraq)", 
         Region != "World total (excluding Iraq)")

```


If you think about `filter()`ing data, so that only certain values of a variable 
are retained, we should remember our friendly `%in%` function! However, if you
think about the code below, you should notice that this code retains only the 
values we wanted to remove! 

```{r}
military_clean %>% 
  filter(Country %in% cont_region$Region)
```


Unfortunately, R doesn't come with a built-in `!%in%` function. However, a
clever way to filter out observations you don't want is with a join. A tool 
tailored just for this scenario is the `anti_join()` function. This function 
will return all of the rows of one dataset **without** a match in another 
dataset. 

__4.__ Use the `anti_join()` function to filter out the `Country` values we 
don't want in the `military_clean` dataset. The `by` argument needs to be 
filled with the name(s) of the variables that the two datasets should be joined 
with. 

**Hint:** To join by different variables on `x` and `y`, use a named vector. For
example, `by = c("a" = "b")` will match `x$a` to `y$b`.

```{r}
military_clean <- military_clean %>% 
  anti_join(cont_region, by = c("Country" = "Region"))
```

</br>

__Part One Answer:__ What regions were not removed from the `military_clean` 
dataset? Use their correct capitalization **and** punctuation! 

- Middle East
- Asia & Oceania
- Central America and the Caribbean
- Sub-Saharan

## Data Organization

The comparison I am interested in looking at the military expenditures of 
different countries across every year in the data. Something like this:  

```{r, echo = FALSE, message = FALSE, warning = FALSE}
military_clean %>%  
  pivot_longer(`1988`:`2019`,
               names_to = "Year", 
               values_to = "Spending") %>% 
  ggplot(aes(y = Year, x = Spending)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(color = "tomato", 
              alpha = 0.2, 
              width = 0.1, 
              height = 0.1) + 
  labs(x = "Spending (% of total Government Spending)")
```

Unfortunately, this requires that each year is it's own column! 

To tidy a dataset like this, we need to gather those columns into a new pair of
variables. To do this process we need three parameters:  

* The set of columns that represent values, not variables. In these data, those
are all the columns from `"1988"` to `"2018"`.  

* The name of the variable to move the column names to. In these data, this 
could be `"Year"`. 

* The name of the variable to move the column values to. In these data, this 
could be labeled `"Spending"`. 

All of these pieces together form the arguments to the `pivot_longer()`
function. 


__4.__ Pivot the cleaned up `military` dataset to a "longer" orientation. Save 
this new "long" version as a new dataset (**do not** overwrite your cleaned 
up dataset)!

```{r}
long_military <- military_clean %>%  
              pivot_longer(`1988`:`2019`,
                           names_to = "Year",
                           values_to = "Spending")

```


## Data Visualization Exploration  

Now that we've transformed the data, let's create a plot to explore the military 
spending across the years.   

__5.__ Create side-by-side boxplots of the military spending for each year. 

**Hint:** Place the `Year` variable on an axis that makes it easy to read the 
labels! 

```{r, warning = FALSE, message = FALSE}

long_military %>% 
  ggplot(aes(x = Spending, y = Year)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(alpha = 0.2, color = "tomato")
  
```


__Part Two Answer:__ What year was the second largest military expenditure? 
What country had this expenditure? 


- Year: 1990
- Country: Kuwait
- Reason: Gulf War

