---
title: " `r emo::ji('speech')` A Few More Data Wrangling Tools </br>"
date: "</br> January 19, 2022"
author: "</br> Dr. Allison Theobold"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---


```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE, 
                      fig.align = "center")

library(xaringanthemer)
library(xaringan)
library(tidyverse)
library(flair)

style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

class: middle

.larger[A few words about `drop_na()`]

--

- Easy tool to remove missing values 

--

- Unilaterally removes **any** row with a missing value for **any** variable

--

- **But** you can specify what columns it should look at for missing values!

---

.larger[Summarizing Frequencies]

`count()` -- counts the values of one or more **categorical** variables

.pull-left[
```{r, eval = FALSE}
starwars |> 
  count(homeworld, sort = TRUE)
```

- `sort` specifies if the resulting tibble should be sorted in descending order
]

--

.pull-right[
```{r, echo = FALSE}
starwars |> 
  count(homeworld, sort = TRUE)
```

]


---

.larger[Unique Groups]

`distinct()` -- selects the unique / distinct rows from a dataset

--

<center>
.large[**Arguments**]
</center>

- `...` -- variables to use when determining uniqueness
  * can use multiple!

--

- `.keep_all` -- decides if all of the columns should be kept
  * `FALSE` is default!


---

.larger[Discritizing Variables]

--

- `if_else()` 
  * Useful when there are two options

--

- `case_when()`
  * Useful when there are three or more options

---

class: middle, center, inverse

.hand[.larger[What if I want to perform the same operation across multiple 
columns?]]

---

class: center, middle

.larger[`across()`]

> makes it easy to apply the same transformation to multiple columns,
> allowing you to use `select()` semantics inside in "data-masking" functions
> like `summarise()` and `mutate()`

--

</br>

.large[`across(.cols = everything(), .fns = NULL, ...)`]

`.cols` -- columns in data to select  
`.fns` -- functions to apply to selected columns  
`...` -- addtional arguments for the function calls (from `.fns`)

---

.larger[Summarizing Multiple Columns]

.pull-left-wide[
```{r, eval = FALSE}
starwars |> 
  summarise(across(height:mass, mean, na.rm = TRUE))

```

</br>
</br>

.large[Conditional Summarizing]
```{r, eval = FALSE}
starwars |> 
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

```
]

.pull-right-narrow[
</br>
```{r, echo = FALSE}
starwars |> 
  summarise(across(height:mass, mean, na.rm = TRUE))

```

</br>
</br>

```{r, echo = FALSE}
starwars |> 
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
```
]

---

.bitlarger[`r emo::ji("heart")` `|>`]

.pull-left[
```{r, eval = FALSE}
starwars |> 
  filter(gender == "feminine", 
         !is.na(homeworld)) |>
  ggplot(aes(x = gender, fill = homeworld)) + 
  geom_bar(position = "dodge") + 
  labs(x = "Gender Expression of Character", 
       fill = "Homeworld")
  
```
]

.pull-right[
```{r, echo = FALSE, fig.width = 6, fig.height = 6}
starwars |> 
  filter(gender == "feminine", 
         !is.na(homeworld)) |>
  ggplot(aes(x = gender, fill = homeworld)) + 
  geom_bar(position = "dodge") +
  labs(x = "Gender Expression of Character", 
       fill = "Homeworld")
  
```
]

