---
title: "Lab 2: Avocado Prices"
subtitle: "Stat 331"
author: "Solutions by Dr. Theobold"
output: rmdformats::readthedown
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width = 12,
                      fig.height = 8,
                      fig.path = 'Figs/',
                      echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      eval = TRUE,
                      include = TRUE)
```


# Instructions

Submit your .html and .Rmd files to the assignment on Canvas.

I advise you to focus particularly on:

* Setting chunk options carefully.

* Making sure you don't print out more output than you need to.

* Making your code readable and nicely formatted. (Next lab, you **will** be 
graded on code style.)

* Thinking through your desired result **before** writing any code.

# Introduction & Set-up

In this lab we're going to be looking at avocado prices. 

The dataset comes to us from kaggle and represents weekly retail scan data. 
It is available to download from Canvas under the name `avocado.csv`.  A
description of the data can be found [here.](https://www.kaggle.com/neuromusic/avocado-prices)

__0.__ Create an R Markdown file, load the data, and declare your package dependencies. 

```{r, message = FALSE}
library(tidyverse)

avocado <- read_csv(here::here("4_tidyr-joins", "Lab", "avocado.csv"))

```

__1.__ Briefly describe the dataset.  What information does it contain?

**Possible Response:** _The data come from weekly retail scales for 
National retail volume and prices of Hass avocados, from 2015 to
2018. There are variables measuring the date of the sale, the average price of 
a single avocado, the region of the sale, the type of avocado (organic 
or conventional), and the size of the avocado._


__2.__ Clean the data in any way you see fit. In particular, look carefully at
the `region` variable.  Notice that the category `"LosAngeles"` is contained in
`"California"`, which is contained in `"West"`, which is contained in 
`"TotalUS"`. Think about how you want to handle this issue in your analysis.


```{r}
# Step 1: Rename the PLU variables to be sizes of avocados! 
avocado <- avocado %>%
  rename(
    Size_Small = `4046`,
    Size_Large = `4225`,
    Size_XL = `4770`)

# Make tibble with the names of the major regions 
major_regions <- tibble(region = c("Plains", "Midsouth", "West", "SouthCentral",
                   "Northeast", "Southeast", "GreatLakes")
                   )
  
# Keep only the observations in the avocado dataset WITH matches in the major_regions
avo_regional <- avocado %>%
  semi_join(major_regions, by = "region")

# Keep only the observations in the avocado dataset WITHOUT matches in the major_regions
avo_metro <- avocado %>%
  anti_join(major_regions, by = "region") %>% 
  ## Filter out "TotalUS" because it isn't a region!
  filter(region != "TotalUS")
```


# Exercises


__1.__ Which major region sold the most organic, small Hass avocados in 2017? 

```{r}
avo_regional %>% 
  filter(type == "organic", 
         year == 2017) %>%
  group_by(region) %>%
  summarize(tot_small_bags = sum(Size_Small)) %>%
  top_n(1, tot_small_bags)
```


__2.__ Use the `separate()` function to split the `Date` variable into year,
month, and day. In which month is the highest volume of avocado sales?

```{r}
avocado %>%
  separate(Date, into = c("Year", "Month", "Day"), sep = "-") %>%
  group_by(Month) %>%
  summarize(tot_sales = sum(`Total Volume`)) %>%
  top_n(1, tot_sales)
```


__3.__ Which metro area regions sold the most total avocados?  Plot side-by-side
boxplots of total volume for only the five regions with the highest averages
for the `Total Volume` variable. 

```{r}
top5reg <- avo_metro %>%
    group_by(region) %>%
    summarize(avgtotvol = mean(`Total Volume`)) %>%
    top_n(5, avgtotvol) %>% 
    pull(region)

avocado %>%
  filter(region %in% top5reg) %>%
  ggplot(aes(y = region, x = `Total Volume`, fill = region)) +
  geom_boxplot() + 
  theme(legend.position = "none")

```


# Reshaping

The following four California regions are in this dataset: `"LosAngeles"`,
`"SanDiego"`, `"Sacramento"`, and `"SanFrancisco"`.  Answer the following
questions about these California regions **only**.  


__1.__ In which regions is the price of organic versus conventional avocados
most different?  Support your answer with a few summary statistics and a 
visualization. 

```{r}

cali <- tibble(region = c("LosAngeles", "SanDiego",
                          "Sacramento", "SanFrancisco")
               )

cali_avo <- avocado %>% 
  semi_join(cali, by = "region")


cali_avo %>% 
  ggplot(aes(x = region, y = AveragePrice, fill = type)) + 
  geom_boxplot() +
  labs(title = "Prices of Avocados in California Regions", 
       x = "CA Region", 
       y = "Average Price of 1 Avocado ($)", 
       fill = "Type of Avocado")


cali_avo %>% 
  group_by(region, type) %>%
  summarize(avg_price = mean(AveragePrice)) %>%
  pivot_wider(names_from = type, 
              values_from = avg_price) %>%
  mutate(price_diff = organic - conventional)
  
```

__2.__  The following plot shows, for all four California regions, the 
proportion of the mean Hass avocado sales that are small, large,
or extra large; conventional vs. organic.  Recreate the plot.

*Hint: This will require restructuring of your data!*


```{r, include = TRUE}
cali_avo %>% 
  group_by(region, type) %>%
  summarize_at(vars(Size_Small, Size_Large, Size_XL), funs(mean)) %>%
  pivot_longer(Size_Small:Size_XL, 
               names_to = "Size", 
               values_to = "Mean_Sold") %>%
  mutate(
    Size = factor(Size, 
                  levels = c("Size_Small", "Size_Large", "Size_XL"),
                  labels = c("Small", "Large", "Extra Large"))
  ) %>%
  ggplot(aes(x = region, y = Mean_Sold, fill = Size)) +
  geom_col(position = "fill") + 
  facet_wrap(~type) + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
  labs(x = "Region of CA", 
       y = "Proportion of Mean Avocados Sold", 
       fill = "Avocado Size")
```


# Challenge 

A joke in the media is that Millenials can't afford houses, because they spend
all their money buying avocado toast.  Let's use this dataset to address that 
claim.

Find or make a dataset with house prices for these four California regions. Join
this dataset with your California avocado dataset.  Use your new joined dataset
to make an argument about the relationship between house prices and avocado prices/sales.  Support your argument with a plot.
