---
title: "Lab 4 Grading Guide"
output: html_document
---

</br>

**If there is no response to a question a score of 0 is awarded.**

**If code outside of the _R for Data Science_ book is used, a citation must be 
included. If no citation is included, the question is awarded a 0.**

</br>

---

# RMarkdown Report 

## Report Formatting

</br>

## Code Formatting 

- 4: Code formatting adheres to the tidyverse styling guide. At a bare minimum, 
the code has the following:
  
  *

- 3: Occassional occurances of code formatting which does not adhere to styling
guide

**Redo if above conditions are not met**

</br>

## Reading in Data

-	4: Reads data in with the `here()` function 
-	Redo: If uses relative or full path to read in the data

</br>

---

# Exercises

## Problem 1

**4**: Must have the following,

1. context of data (**Hass** avocado sales)
2. duration of data (2015 to 2018)
3. description of variables (date of sale, price of avocado, region of sale,
  type of avocado, size of avocado)

**3**: Has components 1 & 2, describes some variables but not all

**Redo if the above conditions are not met**

</br>

## Problem 2

**4**: Must have the following,

1. Uses `rename()`, `mutate()`, or `transmute()` to convert PLU variable names
(`4046`, `4225`, `4770`) to variables related to **size** of avocado

  - `4046` = small
  - `4225` = large
  - `4770` = XL

2. Uses `filter()` to drop `"TotalUS"` from dataset

```{r, eval = FALSE}
avocado <- avocado %>%
  # Rename the PLU variables to be sizes of avocados!
  rename(
    Size_Small = `4046`,
    Size_Large = `4225`,
    Size_XL = `4770`) %>% 
  # Filter out "TotalUS" because it is not a region or city!
  filter(region != "TotalUS")
```

**Redo if not met**


</br>

## Problem 3

**4**: Must have the following,

1. A dataset with the following major regions defined: Plains, Midsouth, 
West, SouthCentral, Northeast, Southeast, GreatLakes"

2. A region-level dataset produced with a `semi_join()`

- 

1. Uses `rename()`, `mutate()`, or `transmute()` to convert PLU variable names
(`4046`, `4225`, `4770`) to variables related to **size** of avocado

  - `4046` = small
  - `4225` = large
  - `4770` = XL

2. Uses `filter()` to drop `"TotalUS"` from dataset

```{r}
# Make tibble with the names of the major regions 
major_regions <- tibble(region = c("Plains",
                                   "Midsouth",
                                   "West",
                                   "SouthCentral",
                                   "Northeast",
                                   "Southeast",
                                   "GreatLakes")
                   )
  
# Keep only the observations in the avocado dataset
# WITH matches in the major_regions
avo_regional <- avocado %>%
  semi_join(major_regions, by = "region")
```
</br>

## Problem 4

</br>

## Problem 5

```{r}
# Make tibble with the names of the states / regions to NOT include
# I'm assuming that "NewYork" is the city and not the state!
states <- tibble(region = c("California",
                            "WestTexNewMexico",
                            "NorthernNewEngland",
                            "SouthCarolina")
                 )

# Keep only the observations in the avocado dataset
# WITHOUT matches in the major_regions
avo_metro <- avocado %>%
  anti_join(major_regions, by = "region") %>% 
  anti_join(states, by = "region")

```


</br>

---

# Reshaping

## Problem 6

</br>

## Problem 7

