---
title: "Lab: The 12 Days of Christmas"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

library(tidyverse)
```

### A Test Dataset

Run the code below to load up a dataset called `xmas` that contains the crucial 
information about the gifts in the song. We will use this dataset to test out
our functions as we work on them.

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```

Note that your functions can - and should! - reference each other.  
That is, don't duplicate code; use earlier, smaller functions inside your larger 
functions.

---

## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day five the
narrator receives "five golden rings", but the entry in the dataset for the gift
on day five simply says "ring".  

Using the skeleton of the `pluralize_gift()` function, complete the code so that 
the function takes a gift and returns the appropriate plural. 

*Hint 1:* Note that the gifts on days six and nine have unusual pluralization.
You may assume that in other datasets, there will be no special cases besides
these types.

*Hint 2:*  The following small example may be useful to you:

```{r, eval = TRUE}
my_names <- c("Kimberly", "Trini", "Jason", "Billy", "Zach", "Tommy")

my_names %>% 
  str_replace("y$", "ies")
```


**Allison's Non-Vectorized Function:** 


```{r}
pluralize_gift <- function(item){
  
  ## Handle goose changing to geese
  if (str_detect(item, "oo")) {
    item <- str_replace(item, "oo", "ee")
  } 
  ## Handle "y" changing to "ies"
  else if (str_detect(item, "y$")) {
    item <- str_replace(item, "y$", "ies")
  } 
  ## Just use an "s" for everything else
  else {
    item <- str_c(item, "s")
  }
  
}
```

### Test Your Function! 

```{r, error = TRUE}
purrr::map_chr(xmas$Gift.Item, pluralize_gift)
```

**Allison's Vectorized Function** 

```{r}
pluralize_gift_vec <- function(item){
  
  plural <- case_when(
     ## Handle goose changing to geese
    str_detect(item, "oo") ~ str_replace(item, "oo", "ee"), 
    ## Handle y changing to ies
    str_detect(item, "y$") ~ str_replace(item, "y$", "ies"), 
    ## Add an s for everything else
    TRUE ~ paste0(item, "s")
  )
  return(plural)
}
```

### Test Your Function

```{r}
pluralize_gift_vec(xmas$Gift.Item)
```

---


## Step Two: Creating sentences

Write a function called `make_phrase()` that takes as input the necessary 
information, and returns a phrase.  


**Allison's Non-Vectorized Function:** 

```{r}
make_phrase <- function(num, num_word, item, verb, adjective, location) {
  
  ## Replace NAs with blank strings
  item <- replace_na(item, "")
  verb <- replace_na(verb, "")
  adjective <- replace_na(adjective, "")
  location <- replace_na(location, "")
  
  ## Figure out if a word starts with a vowel
  vowel_start <- str_sub(item, 1, 1) %>% 
    str_detect("[aeiou]")
  
  ## If the day is larger than 1, the items need pluralized! 
  if (num > 1) {
    item <- pluralize_gift(item)
  } 
  ## If the day is 1 AND the word starts with a vowel, add "an" to the beginning
  else if (vowel_start) {
    num_word <- "an"
  } 
  ## If the day is 1 AND word starts with a consonant, add "a" to the beginning
  else {
    num_word <- "a"
  }
  
  ## Glue all of the pieces together! 
  glue::glue("{num_word} {adjective} {item} {verb} {location}") %>%
    ## Remove extra white space between words 
    str_squish()
  
}
```


### Test Your Function

```{r}
xmas <- xmas %>%
  mutate(
    day.num = as.character(
      english::english(Day)
      ), 
    Full.Phrase = pmap_chr(list(num = Day, num_word = day.num, 
                                item = Gift.Item, verb = Verb, 
                                adjective = Adjective, location = Location), 
                           make_phrase)
  )
```

```{r}
xmas %>% 
  select(Full.Phrase) %>% 
  head()  
```


**Allison's Vectorized Function:** 

```{r}
make_phrase_vec <- function(num, num_word, item, verb, adjective, location) {
  
  ## Replace NAs with blank strings
  item <- replace_na(item, "")
  verb <- replace_na(verb, "")
  adjective <- replace_na(adjective, "")
  location <- replace_na(location, "")
  
  ## Figure out if a word starts with a vowel
  vowel_start <- str_sub(item, 1, 1) %>% 
    str_detect("[aeiou]")
  
  ## If the day is larger than 1, the items need pluralized! 
  item <- if_else(num > 1, pluralize_gift_vec(item), item) 
  
  ## If the word starts with a vowel, add "an" to the beginning
  num_word <- case_when(
    num_word == 1 & vowel_start == TRUE ~ "an",
    num_word == 1 & vowel_start == FALSE ~ "a",
    TRUE ~ num_word)

  ## Glue all of the pieces together! 
  glue::glue("{num_word} {adjective} {item} {verb} {location}") %>%
    str_squish()
}
```


### Test Your Function

```{r}
xmas <- xmas %>%
  mutate(
    day.num = as.character(
      english::english(Day)
      ),
    Full.Phrase = make_phrase_vec(num = Day,
                                  num_word = day.num, 
                                  item = Gift.Item,
                                  verb = Verb, 
                                  adjective = Adjective, 
                                  location = Location)
    )
```

```{r}
xmas %>% 
  select(Full.Phrase) %>% 
  head()  
```

---


## Step Three:  Iteration.

Write a function called `sing_line()` that takes as input:

* A dataset

* A number indicating which day to sing about

* The name of a column in the dataset that contains the phrases for each day.

```{r}
sing_line <- function(data, num, phrase_col) {
  
  ## Set up intro line
  
  num_word <- english::ordinal(num)
  
  intro <- glue::glue("On the {num_word} day of Christmas, my true love gave to me:")
  
  ## Sing gift phrases
  
  phrases <- data %>%
    pull({{phrase_col}})
  
  ## If the number of days is larger than 1
  ## ad an "and" to the second phrase 
  if(num > 1){
    phrases[2] <- paste0(phrases[2], " and")
  }
  
  ## Collapse all of the gifts into a single line 
  ## Start at the day and end at 1
  gift_lines <- str_c(phrases[num:1], collapse = "\n")
  
  ## Put it together
  
  return( 
    glue::glue("{intro} \n{gift_lines}")
  )
  
}
```

### Test Your Function 

```{r}
xmas %>% 
sing_line(num = 2, phrase_col = Full.Phrase)

```


## Step Four: Use Your Functions!

Run the following code to test out your functions! The output should be the 
lyrics for the entire 12 Days of Christmas song. 

```{r}
xmas2 <- xmas %>%
  mutate(
    day.num = as.character(english::english(Day)), 
    Full.Phrase = pmap_chr(list(num = Day, 
                                num_word = day.num, 
                                item = Gift.Item, 
                                verb = Verb, 
                                adjective = Adjective, 
                                location = Location), 
                           make_phrase)
  )

map_chr(1:12, ~sing_line(xmas2, .x, Full.Phrase)) %>%
  str_c(collapse = "\n") %>%
  cat()

```