---
title: "Next Steps -- Selecting Columns"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(xaringan)
library(showtext)
library(fontawesome)
library(tidyverse)
library(flair)

style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

```{r, echo = FALSE, message = FALSE}
cereals <- read_csv("../practice/cereals.csv") %>% 
  mutate(type = as_factor(type), 
         mfr = as_factor(mfr))
```

.large[Selecting]

Verb: `select()`

We **select** which variables we would like to remain in the dataset.

```{r}
select(cereals, 
       name, mfr, calories, cups)
       
```

---

.larger[Selecting a Range of Columns]

You can use the `:` to select a sequence of columns! 

```{r range, eval = FALSE, echo = FALSE}
select(cereals, 
       name:calories)
```

```{r, echo = FALSE}
decorate("range", eval = FALSE) %>% 
  flair(pattern = ":")
```


---

.larger[Omitting Columns]

You can remove columns from the dataset using a `-`. 

```{r omit, eval = FALSE, echo = FALSE}
select(cereals, 
       -rating)
```

```{r, echo = FALSE}
decorate("omit", eval = FALSE) %>% 
  flair(pattern = "-")
```

---

.larger[Reordering]

You can reorder the columns in the dataset inside of `select()`!

```{r}
select(cereals, name, rating, mfr, type, calories, cups, weight, 
       everything())
```


---

.larger[Handy Helpers]

- `:`
- `-`
- `everything()`
- `starts_with()`
- `ends_with()`
- `contains()`

---

.large[A Note on Renaming Columns]

- You can use `select()` to rename, but all columns not included will be deleted! 
  * Using the `rename()` function is easier! 

```{r}
rename(cereals, temp = type)
```



