---
title: "Digging Deeper -- Summarizing Data"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(xaringan)
library(showtext)
library(fontawesome)
library(tidyverse)
library(flair)

style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

```{r, echo = FALSE, message = FALSE}
cereals <- read_csv("../practice/cereals.csv") %>% 
  mutate(type = as_factor(type), 
         mfr = as_factor(mfr))
```


# Summarizing specifics

```{r, echo = FALSE}
decorate('
penguins %>%
  summarize(mean_mass = mean(body_mass_g))
  ') %>% 
  flair("ize(") %>% 
  flair("mean", color = "teal") %>%
  flair("body_mass_g", color = "cornflowerblue")
```

--

* What are the **required** arguments of `summarize`?  

* Are there any optional arguments?

---

# Many variables?  Many summary statistics?
## No problemo

```{r, echo = FALSE}
decorate('
penguins %>%
  summarize(across(c(body_mass_g, flipper_length_mm),
            list(mean, median)))
  ') %>% 
  flair("across") %>%
  flair_rx("(mean)|(median)", color = "teal") %>%
  flair_rx("(body_mass_g)|(flipper_length_mm)", color = "cornflowerblue") %>%
  flair_rx("list", color = "darkred")
```
---
class: center, middle, inverse

# Your Turn!

---
class: center, middle
# In `Practice.Rmd`, find the following:

* Mean, Median, and standard deviation of the sugar content of cereals

* Total calories if I ate one serving of every single cereal in the dataset.
