---
title: "Puzzle Activity: Writing Functions"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```

This activity will ask you to write several small functions, then use them to
unscramble a message.

Many of the functions have been started for you below, but none of them are
complete as-is.

***

__Question 1__ 

Write a function called `divide_and_round()`, which takes a number or vector
of numbers, divides them all by their smallest value, and rounds the results to
the nearest whole number.  

Test it by running the code below the function.

```{r}
divide_and_round <- function(nums){
  my_min <- min(nums, na.rm = TRUE)
  divided <- round( nums / my_min)
  return(divided)
}

test <- c(5:10, NA)

divide_and_round(test)
```


*** 

__Question 2__ 

Write a function called `no_nines_or_twelves()`, which takes a number or
vector of numbers and returns `TRUE` if the numbers are **NOT** divisible
by 9 or 12, and returns `FALSE` otherwise.


```{r}
no_nines_or_twelves <- function(nums){
  check <- !(nums %% 9 == 0 | nums %% 12 == 0)
  return(check)
}

test <- c(3:12*5, NA)

no_nines_or_twelves(test)

```

***

__Question 3__ 

Write a function called `every_other()` which takes a vector and returns every
other value in the vector. 

**Do not** use a `for()` loop to do this! Accomplish this with the `seq_along()` function, bracket subsetting (`[]`), and modulus arithmetic (`%%`). 

Include an optional argument called "start" which lets you choose where to start skipping; that is, if `start = 1`, we return the 1st value, 3rd, 5th, etc. and
if `start = 2`, we return the 2nd, 4th, 6th, etc. 

```{r}

every_other <- function(vec, start = 1){
  
  if(start == 2){
    return(
      vec[(seq_along(vec) + 1) %% 2 > 0]
    )
  } else if(start == 1) {
    return(
      vec[seq_along(vec) %% 2 > 0]
      ) 
  }
}

test <- c(1:10)

every_other(test)
every_other(test, start = 2)
```

***

__Question 4__ 

Write a function called `shorten()` which takes a vector, drops the values 
**before** the cumulative sum is greater than a provided number, and returns the
remaining values.

**Do not** use a `while()` loop to do this! Accomplish this with the ``cumsum()` function and bracketing.

Be sure to test your function!

```{r}
shorten <- function(x, sum){
  vec_sum <- cumsum(x) 
  return(
    x[vec_sum > sum]
  )
}


test <- 1:100 
shorten(test, 350)
```

***

__Question 5__ 

Once you have written your four functions correctly, copy-paste and run the
following code:

```{r}

my_vec <- c(39, 1.87, 48, 11, 8, 45, 21, 5, 12, 33, 9, 11, 108, 4, 18, 5, 16,
            17, 8, 48, 27, 24, 4, 22, 12, 27, 23, 46, 42, 35, 15, 34, 36, 26,
            18, 10, 18.21, 72.04, 36.9, 41.81, 29, 89.75, 34.03, 20.18, 48.74,
            15.76, 31.86, 83.6, 43.55, 39.99, 23.55, 8.54, 24.71, 22.02, 9.71,
            62.14, 35.46, 16.61, 15.66, 21.29, 30.52, 201.07, 45.81, 7.85,
            30.13, 34.14, 22.62, 10.2, 6.02, 30.12, 10.64, 31.72, 24.57, 14.43,
            43.37, 89.93, 44.72, 51.32, 13.62, 45.56, 22.96, 7.05, 29.99, 41.38,
            26.59, 23.04, 19.82, 50.73, 39.56, 43.79, 30.22, 85.85, 5.78, 78.85,
            29.52, 66.27, 44.06, 27.28, 24.43, 64.32, 3.35, 67.45, 46.72, 48.44,
            48.65, 33.3, 40.28, 19.04)

my_vec <- every_other(my_vec, start = 2)
my_vec <- divide_and_round(my_vec)
my_vec <- every_other(my_vec, start = 1)
my_vec <- shorten(my_vec, 350)
my_vec <- my_vec[no_nines_or_twelves(my_vec)]

my_vec <- sort(my_vec)

my_vec

```

If you have done everything correctly, your final vector will be six numbers
long.  Google these six numbers to find a TV show as your final answer.
