---
title: "Iteration with the `apply()` family"
author: Stat 331
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---


```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE, 
                      fig.align = "center")

options(htmltools.dir.version = FALSE)

library(xaringanthemer)
library(xaringan)
library(tidyverse)
library(flair)
library(palmerpenguins)

style_duo_accent(
  primary_color      = "#0F4C81", # pantone classic blue
  secondary_color    = "#B6CADA", # pantone baby blue
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  text_font_size     = "30px"
)
```

# Better: The `apply` functions

```{r}

sapply(99:95, sing_about_pop)

```
---

# Better: The `apply` functions

- Pay attention to what type of input they *take* and what type of output they *give*

- `sapply` takes a vector, returns a vector

- `lapply` takes a list, returns a list

- `apply` takes a dataset and a dimension (1 = row, 2 = col), returns a vector or list

---

# What if I want some options in my function?

```{r}
sing_about_drinks <- function(num, drink) {
  
  
    
  glue("{num} bottles of {drink} on the wall, {num} bottles of {drink}.
       Take one down, pass it around, {num - 1} bottles of {drink} on the wall.")
  
}
```

---
# `sapply` has trouble!

```{r, error = TRUE}

sapply(99:95, sing_about_drinks(drink = "beer"))

```
---

# You can kind of get around this...

(This is called a **headless** or **annonymous** function)

```{r}

sapply(99:95, function(x) sing_about_drinks(x, drink = "beer"))

```
