---
title: "Lab 1 Solutions"
author: "Dr. Theobold"
output: html_document
---

# RMarkdown Output Options

```{r setup}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center", 
                      messages = FALSE)
```

## Setup

```{r load-libraries-data, include = FALSE}
library(tidyverse)
```
    
## Car Dataset
    
This dataset contains information about the qualities of certain car models made
in 1999 and 2008.
    
It has 234 rows and 11 variables.

## Including Plots

```{r pressure, echo=FALSE}
plot(pressure, 
     xlab = "Temperature (deg C)", 
     ylab = "Pressure (mm)")
```

## Exploring the Cars Dataset

(The plots below are only one option out of many possibilities, of course.)

### Step 1: Extract the Distances

```{r}
dist <- cars$dist
```

### Step 2: Plot the Distribution of Distances

*Distributions* of a single numerical variable are plotted using histograms or 
density plots

**Option 1**: Using the `hist()` function 

```{r}
hist(dist, xlab = "Stopping Distance (ft)")
```

**Option 2**: Using `ggplot()`

```{r}
dist %>% 
  as_tibble() %>% 
  ggplot(mapping = aes(x = value)) + 
  geom_histogram(binwidth = 10) + 
  labs(x = "Stopping Distance (ft)")
```

## Step 3: Stopping distance and Speed

**Option 1**: Using the `plot()` function 

```{r}
plot(x = cars$speed,
     y = cars$dist, 
     xlab = "Stopping Distance (ft)", 
     ylab = "Speed (mph)")
```

**Option 2**: Using `ggplot()`

```{r}
cars %>% 
  ggplot(mapping = aes(x = speed, y = dist)) + 
  geom_point() + 
  labs(x = "Stopping Distance (ft)", 
       y = "Speed (mph)")
```

