---
title: "Desired Plot Modifications"
author: "Compiled by STAT 331"
date: "Winter 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Histogram

```{r cars}
cars %>% 
  ggplot(mapping = aes(x = dist)) + 
  geom_histogram(binwidth = 10, aes(y = ..density..)) + 
  geom_density() +
  labs(x = "Stopping Distance (ft)", 
       title = "1920s Cars")
```

- The `binwidth` argument allows for you to specify how large each bin should be

- To add a density layer on top of the histogram, the y-axis of the histogram
needs to be plotted as proportions not a counts. To do this, you override the 
y aesthetic (`y = ..density..`). 


## Scatterplot

```{r, echo = FALSE}
cars %>% 
  ggplot(mapping = aes(x = speed, y = dist)) + 
  geom_point(color = "blue", shape = "circle", size = 2) + 
  geom_smooth(method = "lm") +
  labs(x = "Stopping Distance (ft)", 
       y = "Speed (mph)", 
       title = "Relationship between stopping distance and speed \nfor 1920s Automobiles") +
  scale_y_continuous(limits = c(0, 30), breaks = seq(0, 30, by = 5)) +
  theme_bw()


```

- If you insert a `\n` in the plot title, `labs()` interprets that as a command 
to start a new line

- The `shape` argument of `geom_point()` allows for you to specify what type of 
plotting character should be used. The character can be specified with either a 
number (http://www.sthda.com/english/wiki/ggplot2-point-shapes) or a string
associated with that number (see <datanovia.com/en/blog/ggplot-point-shapes-best-tips/> 
for the names associated with each number).

- The `size` argument of `geom_point()` allows for you to specify what size the
plotting characters should be. 

- You can add a line on top of the points by adding a `geom_smooth()` layer!
There are different methods for a smoother, but the `"lm"` method stands for a 
linear smoother.

- 



