---
title: "Reproducibility: Directories and Deliverables"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
```

The theme of this lesson is good management of your files and data.

You will learn how to:

-   Identify **folders** and **paths**.

-   Create and use **R Projects**.

-   **Load data** into R from files.

-   Deal with package challenges.

-   Create R Markdown or Quarto documents with good **reproducible
    principles**.

------------------------------------------------------------------------

`r emo::ji("play button")` Watch Videos: 15 min

`r emo::ji("book")` Readings: 15 minutes

`r emo::ji("computer")` Activities: 10 min

`r emo::ji("check")` Check-ins: 2

------------------------------------------------------------------------

## Principles of Reproducibility

`r emo::ji("play button")` [Required Video: Directories, Paths, Projects
(15 minutes)](http://somup.com/c3QDczUaCa)

`r emo::ji("book")` [Required Reading: Workflow and
Projects](https://r4ds.had.co.nz/workflow-projects.html)

### `r emo::ji("question")` Extra Resources

- [Software Carpentry -- Project Managment with RStudio](http://swcarpentry.github.io/r-novice-gapminder/02-project-intro/index.html)

------------------------------------------------------------------------

`r emo::ji("check")` Check-in: RStudio Projects

Take a screenshot of your class directory, showing the **R Project** and
the folder organization, including a **Week 1** or **Preview Activity**
folder.

------------------------------------------------------------------------

## Reproducible Deliverables

We will be using Quarto documents in this course. To compile a Quarto document,
you first need to download the Quarto software. You can do that [here](). 

Quarto was released less than six months ago, so the *R for Data Science* 
textbook has not had time to write a chapter specifically dedicated to these
documents. So, instead we are going to read through elements of the Quarto 
documentation. 

### Introduction to Quarto

`r emo::ji("book")` [Required Reading: Intro to Quarto](https://quarto.org/docs/get-started/hello/rstudio.html)

`r emo::ji("book")` [Required Reading: Computations in Quarto](https://quarto.org/docs/get-started/computations/rstudio.html)

`r emo::ji("book")` [Required Reading: Authoring & Formatting Quarto Documents](https://quarto.org/docs/get-started/authoring/rstudio.html)

- In this class, we will only render HTML documents, so you can ignore the 
sections describing PDF and Word files. 

------------------------------------------------------------------------

# `r emo::ji("question")` Extra Resources

- You can find the entire list of options you can use to format your HTML file 
with Quarto [here](https://quarto.org/docs/output-formats/html-basics.html). 

- Poke around the gallery of cool HTML documents rendered with Quarto:
  * [Interactive document published to the web](https://quarto-dev.github.io/quarto-gallery/articles/html/html.html)
  * [Advanced Layout with HTML](https://quarto-dev.github.io/quarto-gallery/page-layout/tufte.html)
  
