---
title: "Reproducibility: Directories and Deliverables"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
```

The theme of this lesson is good management of your files and data.

You will learn how to:

* Identify **folders** and **paths**.

* Create and use **R Projects**.

* **Load data** into R from files.

* Deal with package challenges.

* Create R Markdown or Quarto documents with good **reproducible principles**.

---

`r emo::ji("play button")` Watch Videos: 25 min

`r emo::ji("book")` Readings: 15 minutes 

`r emo::ji("computer")` Activities: 10 min 

`r emo::ji("check")` Check-ins: 2

---

## Principles of Reproducibility

`r emo::ji("play button")` [Required Video: Directories, Paths, Projects (17 minutes)]()

`r emo::ji("book")` [Required Reading: Workflow and Projects](https://r4ds.had.co.nz/workflow-projects.html) 

---

## Reproducible Deliverables

### Introduction to R Markdown

`r emo::ji("book")` [Required Reading: Intro to R Markdown](https://r4ds.had.co.nz/r-markdown.html)

### Introduction to Quarto

`r emo::ji("book")` [Required Reading: Intro to Quarto](https://quarto.org/docs/get-started/hello/rstudio.html)

### Comparison of Quarto & R Markdown

`r emo::ji("play button")` [Required Video: What's the difference between Quarto and R Markdown? (17 minutes)]()

---

# `r emo::ji("question")` Extra Resources

* [Software Carpentry -- Project Managment with RStudio](http://swcarpentry.github.io/r-novice-gapminder/02-project-intro/index.html)

---

`r emo::ji("check")` Check-in: RStudio Projects

Take a screenshot of your class directory, showing the **R Project** and the
folder organization, including a **Week 1** or **Practice Activity** folder.

---

# `r emo::ji("question")` Extra Resources

* [Click Here](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
to download a cheatsheet for R Markdown.

* [This website](https://holtzy.github.io/Pimp-my-rmd/) has some tips for making
your R Markdown documents more beautiful.

* [This website](https://quarto.org/docs/gallery/) has a gallery of different
types of products you can make with Quarto. I especially like the [HTML with
advanced layout option](https://quarto-dev.github.io/quarto-gallery/page-layout/tufte.html), 
since I really like adding comments and plots in the margins of my document. 

